---
title: "MP-Game-5: Multiplayer Boolean Concept Learning (Summurization Format)"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(reshape2)
library(purrr)
library(jsonlite)
library(tidyr) 
library(dplyr)
library(tidyboot)
library(ggplot2)
library(scales)
library(stringr)
library(digest)
```

```{r}
#-----------------------
# Reading Mp-Game-3 Data
#-----------------------
# Load Training Trial Responses
temp <- list.files(
  "../../../data/mp-game-3/experiment/train_trials",
  pattern="*.csv",
  full.names=TRUE
)
train_trials_rbr <- do.call(rbind, lapply(temp, read.csv))

# Load Test Trial Responses
temp <- list.files(
  "../../../data/mp-game-3/experiment/test_trials",
  pattern="*.csv",
  full.names=TRUE
)
test_trials_rbr <- do.call(rbind, lapply(temp, read.csv))

# Load Training Summary Stats
temp <- list.files(
  "../../../data/mp-game-3/experiment/train_summary_stats",
  pattern="*.csv",
  full.names=TRUE
)
train_stats_rbr <- do.call(rbind, lapply(temp, read.csv))

# Load Test Summary Stats
temp <- list.files(
  "../../../data/mp-game-3/experiment/test_summary_stats",
  pattern="*.csv",
  full.names=TRUE
)
test_stats_rbr <- do.call(rbind, lapply(temp, read.csv))

# Load Test Summary Stats
temp <- list.files(
  "../../../data/mp-game-3/experiment/test_summary_stats",
  pattern="*.csv",
  full.names=TRUE
)
test_stats_rbr <- do.call(rbind, lapply(temp, read.csv))
```

```{r}
#-----------------------
# Reading Mp-Game-5 Data
#-----------------------
# Load Training Trial Responses
temp <- list.files(
  "../../../data/mp-game-5/logTrain",
  pattern="*.csv",
  full.names=TRUE
)
train_trials_s <- do.call(rbind, lapply(temp, read.csv, sep=""))

# Load Test Trial Responses
temp <- list.files(
  "../../../data/mp-game-5/logTest",
  pattern="*.csv",
  full.names=TRUE
)
test_trials_s <- do.call(rbind, lapply(temp, read.csv, sep=""))

# Load Test Summary Stats
temp <- list.files(
    "../../../data/mp-game-5/logScores",
  pattern="*.csv",
  full.names=TRUE
)
test_stats_s <- do.call(rbind, lapply(temp, read.csv, sep=""))

# Load Chat Messages
temp <- list.files(
    "../../../data/mp-game-5/chatMessage",
  pattern="*.tsv",
  full.names=TRUE
)
chat_messages <- do.call(rbind, lapply(temp, readr::read_tsv)) %>%
  group_by(gameid) %>%
  mutate(num_dialogues = n()) %>%
  filter(num_dialogues > 10)
```


```{r}
rule_indices <- c(0, 2, 4, 6, 8)

#--------------------------------------
# Accuracy Calculations Round-By-Round
# -------------------------------------
test_acc_teacher_rbr <- test_stats_rbr %>%
  filter(role == "explorer") %>%
  mutate(correct = hits + correct_rejections) %>%
  mutate(incorrect = misses + false_alarms) %>%
  select(rule_idx, correct, incorrect) %>%
  filter(rule_idx %in% rule_indices) %>%
  group_by(rule_idx) %>% 
  summarise_all(sum) %>%
  mutate(acc = correct/(correct + incorrect)) %>%
  mutate(game = "round")
test_acc_teacher_rbr$rule_idx <- as.factor(test_acc_teacher_rbr$rule_idx)
test_acc_teacher_rbr$game <- as.factor(test_acc_teacher_rbr$game)

test_acc_student_rbr <- test_stats_rbr %>%
  filter(role == "student") %>%
  mutate(correct = hits + correct_rejections) %>%
  mutate(incorrect = misses + false_alarms) %>%
  select(rule_idx, incorrect, correct) %>%
  filter(rule_idx %in% rule_indices) %>%
  group_by(rule_idx) %>% 
  summarise_all(sum) %>%
  mutate(acc = correct/(correct + incorrect)) %>%
  mutate(game = "round")
test_acc_student_rbr$rule_idx <- as.factor(test_acc_student_rbr$rule_idx)
test_acc_student_rbr$game <- as.factor(test_acc_student_rbr$game)

#--------------------------------------
# Accuracy Calculations Summurization
# -------------------------------------
test_acc_teacher_s <- test_stats_s %>%
  filter(role == "explorer") %>%
  mutate(correct = hits + correct_rejections) %>%
  mutate(incorrect = misses + false_alarms) %>%
  select(rule_idx, correct, incorrect) %>%
  filter(rule_idx %in% rule_indices) %>%
  group_by(rule_idx) %>% 
  summarise_all(sum) %>%
  mutate(acc = correct/(correct + incorrect)) %>%
  mutate(game = "summurization")
test_acc_teacher_s$game <- as.factor(test_acc_teacher_s$game)
test_acc_teacher_s$rule_idx <- as.factor(test_acc_teacher_s$rule_idx)

test_acc_student_s <- test_stats_s %>%
  filter(role == "student") %>%
  mutate(correct = hits + correct_rejections) %>%
  mutate(incorrect = misses + false_alarms) %>%
  select(rule_idx, incorrect, correct) %>%
  filter(rule_idx %in% rule_indices) %>%
  group_by(rule_idx) %>% 
  summarise_all(sum)%>%
  mutate(acc = correct/(correct + incorrect)) %>%
  mutate(game = "summurization")
test_acc_student_s$game <- as.factor(test_acc_student_s$game)
test_acc_student_s$rule_idx <- as.factor(test_acc_student_s$rule_idx)


#----------------------------
# Accuracy Calculations Joint
# ---------------------------
test_acc_teacher_combined = bind_cols(
  test_acc_teacher_rbr %>%
    select(rule_idx, acc) %>%
    rename(round_by_round_acc = acc),
  test_acc_teacher_s %>%
    select(rule_idx, acc) %>%
    rename(summurization_acc = acc)
  ) %>%
  select(rule_idx, round_by_round_acc, summurization_acc)

test_acc_student_combined = bind_cols(
  test_acc_student_rbr %>%
    select(rule_idx, acc) %>%
    rename(round_by_round_acc = acc),
  test_acc_student_s %>%
    select(rule_idx, acc) %>%
    rename(summurization_acc = acc)
  ) %>%
  select(rule_idx, round_by_round_acc, summurization_acc)
```

## Intro
Our goal is to ollect cheaper data for cultural ratchet experiment. In the spring, we ran a pilot experiment, with a "round-by-round" format, where players only learned a single concept. In fall, we ran a pilot experiment with a "summurization" format, where palyers played multiple concepts. For spring pilot (round-by-round), we had 5 concepts, each with two different lists of stimuli. For fall pilot (summurization), we had the same 5 concepts and selected one of two lists of stimuli, from spring.


## Cost Comparison
Spring Pilot Cost Per Round (Round-By-Round-Format):

  $122.40 / 36 = $3.40 / round
  
Fall Pilot Cost Per Round (Summurization-Format):

  ($(1.25 * 20 + .50 * 20) + $11.60)) / 40 = $1.165 / round

If we were maximally efficient, the Fall Pilot would be $.892 / round. We had two games crash midway because of players disconnecting; so our cost for the pilot was slightly higher than expected according to the data we collected.

## Experiment Details (Critters)
* 81 total possible critters: ~50 training, ~31 test
* 4 axes of variability:
    + Critter Type (Bug, Fish, Bird)
    + Primary Color (Blue, Green, Orange)
    + Secondary Color (Red, Yellow, Purple)
    + Size (Small, Medium, Large)


## Experiment Details (Concepts)
* Rule Idx 0: Primary Color == Orange ("Orange things")
* Rule Idx 2: Critter Type == Fish && Primary Color == Blue ("Blue fish")
* Rule Idx 4: Primary Color == Orange && Secondary Color == Purple ("Purple and orange things")
* Rule Idx 6: Critter Type == Bug || Secondary Color == Yellow ("Bugs, or yellow things")
* Rule Idx 8: Critter Type == Bird || Primary Color == Green ("Birds, or green things")


# Round vs. Summurization

## Teacher Performance (Round vs. Summurization Acc.)
```{r}
ggplot(test_acc_teacher_combined, aes(x=round_by_round_acc, y=summurization_acc, color=rule_idx)) +
  geom_point() +
  xlim(0, 1) +
  ylim(0, 1)
```


## Student Performance (Round vs. Summurization Acc.)
```{r}
ggplot(test_acc_student_combined, aes(x=round_by_round_acc, y=summurization_acc, color=rule_idx)) +
  geom_point() +
  xlim(0, 1) +
  ylim(0, 1)
```

```{r}
rule_indices <- c(0, 2, 4, 6, 8)

#-----------------------------------------------------------------
# Hits, Misses, False Alarms, Correct Rejections -- Round By Round
# ----------------------------------------------------------------
test_stats_teacher_rbr <- test_stats_rbr %>%
  filter(role == "explorer") %>%
  select(rule_idx, hits, misses, correct_rejections, false_alarms) %>%
  filter(rule_idx %in% rule_indices) %>%
  group_by(rule_idx) %>% 
  summarise_all(sum)
summed <- rowSums(test_stats_teacher_rbr[, c("hits", "misses", "correct_rejections", "false_alarms")])
test_stats_teacher_rbr_normalized <- test_stats_teacher_rbr %>%
  mutate(hits = hits / summed) %>%
  mutate(misses = misses / summed) %>%
  mutate(correct_rejections = correct_rejections / summed) %>%
  mutate(false_alarms = false_alarms / summed)

test_stats_student_rbr <- test_stats_rbr %>%
  filter(role == "student") %>%
  select(rule_idx, hits, misses, correct_rejections, false_alarms) %>%
  filter(rule_idx %in% rule_indices) %>%
  group_by(rule_idx) %>% 
  summarise_all(sum)
summed <- rowSums(test_stats_student_rbr[, c("hits", "misses", "correct_rejections", "false_alarms")])
test_stats_student_rbr_normalized <- test_stats_student_rbr %>%
  mutate(hits = hits / summed) %>%
  mutate(misses = misses / summed) %>%
  mutate(correct_rejections = correct_rejections / summed) %>%
  mutate(false_alarms = false_alarms / summed)

#-----------------------------------------------------------------
# Hits, Misses, False Alarms, Correct Rejections -- Summurization
# ----------------------------------------------------------------
test_stats_teacher_s <- test_stats_s %>%
  filter(role == "explorer") %>%
  select(rule_idx, hits, misses, correct_rejections, false_alarms) %>%
  filter(rule_idx %in% rule_indices) %>%
  group_by(rule_idx) %>% 
  summarise_all(sum)
summed <- rowSums(test_stats_teacher_s[, c("hits", "misses", "correct_rejections", "false_alarms")])
test_stats_teacher_s_normalized <- test_stats_teacher_s %>%
  mutate(hits = hits / summed) %>%
  mutate(misses = misses / summed) %>%
  mutate(correct_rejections = correct_rejections / summed) %>%
  mutate(false_alarms = false_alarms / summed)

test_stats_student_s <- test_stats_s %>%
  filter(role == "student") %>%
  select(rule_idx, hits, misses, correct_rejections, false_alarms) %>%
  filter(rule_idx %in% rule_indices) %>%
  group_by(rule_idx) %>% 
  summarise_all(sum)
summed <- rowSums(test_stats_student_s[, c("hits", "misses", "correct_rejections", "false_alarms")])
test_stats_student_s_normalized <- test_stats_student_s %>%
  mutate(hits = hits / summed) %>%
  mutate(misses = misses / summed) %>%
  mutate(correct_rejections = correct_rejections / summed) %>%
  mutate(false_alarms = false_alarms / summed)

#--------------------------------------------------------
# Hits, Misses, False Alarms, Correct Rejections -- Joint
# -------------------------------------------------------
test_stats_teacher_combined_normalized = bind_rows(
  test_stats_teacher_s_normalized %>%
    gather(key = "stat", value = "val", -rule_idx) %>%
    arrange(rule_idx) %>%
    select(rule_idx, val, stat) %>%
    mutate(game="s"),
  test_stats_teacher_rbr_normalized %>%
    gather(key = "stat", value = "val", -rule_idx) %>%
    arrange(rule_idx) %>%
    select(rule_idx, val, stat) %>%
    mutate(game="r")
  )
test_stats_teacher_combined_normalized$game <- as.factor(test_stats_teacher_combined_normalized$game)
test_stats_teacher_combined_normalized$rule_idx <- as.factor(test_stats_teacher_combined_normalized$rule_idx)

test_stats_student_combined_normalized = bind_rows(
  test_stats_student_s_normalized %>%
    gather(key = "stat", value = "val", -rule_idx) %>%
    arrange(rule_idx) %>%
    select(rule_idx, val, stat) %>%
    mutate(game="s"),
  test_stats_student_rbr_normalized %>%
    gather(key = "stat", value = "val", -rule_idx) %>%
    arrange(rule_idx) %>%
    select(rule_idx, val, stat) %>%
    mutate(game="r")
  )
test_stats_student_combined_normalized$game <- as.factor(test_stats_student_combined_normalized$game)
test_stats_student_combined_normalized$rule_idx <- as.factor(test_stats_student_combined_normalized$rule_idx)
```


## Teacher Performance (Round vs. Summurization)
```{r}
ggplot(test_stats_teacher_combined_normalized, aes(x = game, y = val, fill = stat)) + 
  geom_bar(stat = 'identity', position = 'stack') + facet_grid(~ rule_idx)
```


## Student Performance (Round vs. Summurization)
```{r}
ggplot(test_stats_student_combined_normalized, aes(x = game, y = val, fill = stat)) + 
  geom_bar(stat = 'identity', position = 'stack') + facet_grid(~ rule_idx)
```

```{r}
test_stats_s_unique = test_stats_s %>%
  mutate(role = as.character(role)) %>%
  mutate(role = replace(role, role == "explorer", "T")) %>%
  mutate(role = replace(role, role == "student", "S")) %>%
  mutate(id = strtoi(str_sub(test_stats_s$gameid, 0, 4)) + rule_idx) %>%
  select(id, role, rule_idx, hits, misses, correct_rejections, false_alarms) %>%
  gather(key = "stat", value = "val", 4:7) %>%
  arrange(id) %>%
  select(id, role, val, stat)
```

# Summurization Game Analysis

## Student Performance Vs. Teacher Performance
```{r}
rule_indices <- c(0, 2, 4, 6, 8, 10, 11, 12, 13, 14)

#--------------------------------------
# Accuracy Calculations Summurization
# -------------------------------------
test_teacher <- test_stats_s %>%
  filter(role == "explorer") %>%
  mutate(correct = hits + correct_rejections) %>%
  mutate(incorrect = misses + false_alarms) %>%
  select(gameid, rule_idx, correct, incorrect) %>%
  filter(rule_idx %in% rule_indices) %>%
  mutate(teacher_acc = correct / (correct + incorrect))

test_teacher_summary <- test_teacher %>%
  group_by(rule_idx) %>%
  summarize(
    avg_teacher_acc = mean(teacher_acc),
    std_teacher_acc = sd(teacher_acc),
  ) %>%
  select(rule_idx, avg_teacher_acc, std_teacher_acc)

test_student <- test_stats_s %>%
  filter(role == "student") %>%
  mutate(correct = hits + correct_rejections) %>%
  mutate(incorrect = misses + false_alarms) %>%
  select(gameid, rule_idx, incorrect, correct) %>%
  mutate(student_acc = correct / (correct + incorrect))

test_student_summary <- test_student %>%
  group_by(rule_idx) %>%
  summarize(
    avg_student_acc = mean(student_acc),
    std_student_acc = sd(student_acc),
  ) %>%
  select(rule_idx, avg_student_acc, std_student_acc)

test_summary <- bind_cols(test_teacher_summary, test_student_summary) %>%
  select(rule_idx, avg_student_acc, avg_teacher_acc, std_student_acc, std_teacher_acc)
test_summary$rule_idx <- as.factor(test_summary$rule_idx)

ggplot(test_summary, aes(x=avg_student_acc, y=avg_teacher_acc, color=rule_idx, label=rule_idx)) +
  geom_point() +
  xlim(0, 1.1) +
  ylim(0, 1.1) +
  geom_text(hjust = 0.01, nudge_x = 0.03, show.legend = FALSE) +
  geom_errorbar(test_summary, mapping=aes(x=avg_student_acc, ymin=avg_teacher_acc - std_teacher_acc, ymax=avg_teacher_acc + std_teacher_acc)) +
  geom_errorbarh(test_summary, mapping=aes(y=avg_teacher_acc, xmin=avg_student_acc - std_student_acc, xmax=avg_student_acc + std_student_acc)) +
  ggtitle("Avg. Student Acc vs. Avg. Teacher Acc.") +
  xlab('Avg. Student Acc.') +
  ylab('Avg. Teacher Acc.')
```



## Student versus Teacher Performance -- Pairs (Summurization)
```{r}
ggplot(test_stats_s_unique, aes(x = reorder(role, desc(role)), y = val, fill = stat)) + 
  geom_bar(stat = 'identity', position = 'stack') + facet_grid(~ id)
```

```{r}
test_stats_s_concept = test_stats_s %>%
  mutate(role = as.character(role)) %>%
  mutate(role = replace(role, role == "explorer", "T")) %>%
  mutate(role = replace(role, role == "student", "S")) %>%
  select(role, rule_idx, hits, misses, correct_rejections, false_alarms) %>%
  group_by(rule_idx, role) %>%
  summarize(hits = sum(hits), misses = sum(misses), correct_rejections=sum(correct_rejections), false_alarms=sum(false_alarms)) %>%
  gather(key = "stat", value = "val", 3:6) %>%
  arrange(rule_idx) %>%
  select(rule_idx, role, val, stat)
```
## Student versus Teacher Performance -- Concepts (Summurization)
```{r}
ggplot(test_stats_s_concept, aes(x = reorder(role, desc(role)), y = val, fill = stat)) + 
  geom_bar(stat = 'identity', position = 'stack') + facet_grid(~ rule_idx) 
```

```{r}
test_stats_s_rule_type = test_stats_s %>%
  mutate(role = as.character(role)) %>%
  mutate(role = replace(role, role == "explorer", "T")) %>%
  mutate(role = replace(role, role == "student", "S")) %>%
  select(role, rule_type, hits, misses, correct_rejections, false_alarms) %>%
  group_by(rule_type, role) %>%
  summarize(hits = sum(hits), misses = sum(misses), correct_rejections=sum(correct_rejections), false_alarms=sum(false_alarms)) %>%
  gather(key = "stat", value = "val", 3:6) %>%
  select(rule_type, role, val, stat)
```

## Student versus Teacher Performance -- Concept Types (Summurization)
```{r}
ggplot(test_stats_s_rule_type, aes(x = reorder(role, desc(role)), y = val, fill = stat)) + 
  geom_bar(stat = 'identity', position = 'stack') + facet_grid(~ rule_type) 
```

## Time Comparisons (Summarization)
```{r}
train_times <- train_trials_s %>%
    select(rule_idx, total_time) %>%
    filter(rule_idx %in% rule_indices) %>%
    group_by(rule_idx) %>% 
    summarise_all(mean) %>%
    mutate(role = "T") %>%
    rename(train_time = total_time)

test_times <- bind_rows(
  test_stats_s %>%
    filter(role == "explorer") %>%
    select(rule_idx, total_time) %>%
    filter(rule_idx %in% rule_indices) %>%
    group_by(rule_idx) %>% 
    summarise_all(mean) %>%
    mutate(role = "T"),
  
  test_stats_s %>%
    filter(role == "student") %>%
    select(rule_idx, total_time) %>%
    filter(rule_idx %in% rule_indices) %>%
    group_by(rule_idx) %>% 
    summarise_all(mean) %>%
    mutate(role = "S")
) %>%
  rename(test_time = total_time)

chat_times <- chat_messages %>%
    filter(rule_idx %in% rule_indices) %>%
    select(rule_idx, gameid, reactionTime) %>%
    group_by(rule_idx, gameid) %>%
    summarize(reactionTime = sum(reactionTime)) %>%
    group_by(rule_idx) %>%
    summarize(reactionTime = mean(reactionTime) / 1000) %>%
    rename(chat_time = reactionTime)

train_chat_times <- bind_cols(train_times, chat_times) %>%
  select(rule_idx, train_time, chat_time, role) %>%
  gather(key = "stat", value = "val", 2:3) %>%
  select(rule_idx, role, val, stat)
```


```{r}
ggplot(train_chat_times, aes(x = reorder(stat, desc(stat)), y = val)) + 
  geom_bar(stat = 'identity') + facet_grid(~ rule_idx)
```

For Simple Rules 0, 2, 4 (Single Feature & Conjunction) we see that there isn't a wide variance in terms of train time and chat time; but this rift becomes noticeable when we have Harder Rules 6, 8 (Disjunction).

# Corpus Analysis

## Breakdown of Dialogue Statements (Categories)
```{r}
chat_messages_count = chat_messages %>%
  mutate(role = as.character(role)) %>%
  mutate(role = replace(role, role == "explorer", "T")) %>%
  mutate(role = replace(role, role == "student", "S")) %>%
  select(role, messageType) %>%
  group_by(role) %>%
  summarize(
    informative = sum(messageType == 'Informative'),
    follow_up = sum(messageType == 'Follow-Up'),
    social = sum(messageType == 'Social'),
    misc = sum(messageType == 'Misc'),
    confirmation = sum(messageType == 'Confirmation'),
  ) %>%
  select(role, informative, follow_up, social, misc, confirmation) %>%
  gather(key = "stat", value = "val", 2:6) %>%
  select(role, val, stat) %>%
  group_by(role) %>%
  mutate(val_percentage_of_role = val/sum(val))

```
```{r}
ggplot(chat_messages_count, aes(x = reorder(role, desc(role)), y = val_percentage_of_role, fill = stat)) + 
  geom_bar(stat = 'identity', position = 'stack') + ggtitle("% of dialogue messages produced by each player") + xlab('Role') + ylab('% of player\'s messages')
```

```{r}
ggplot(chat_messages_count, aes(x = reorder(role, desc(role)), y = val, fill = stat)) + 
  geom_bar(stat = 'identity', position = 'stack') + ggtitle("# of messages produced by each player") +
  xlab('Role') + ylab('# of player\'s messages')
```



```{r}
ggplot(
    chat_messages_count %>%
      select(val, stat) %>%
      group_by(stat) %>%
      summarize(val = sum(val)),
    aes(x = stat, y = val)
  ) + geom_bar(stat = 'identity') + ggtitle("# of Messages, Across Type") +
    xlab('Message Type') + ylab('# of Messages Produced')
```



```{r}
chat_messages_word_count = chat_messages %>%
  mutate(role = as.character(role)) %>%
  mutate(role = replace(role, role == "explorer", "T")) %>%
  mutate(role = replace(role, role == "student", "S")) %>%
  select(role, messageType, text) %>%
  group_by(role, messageType) %>%
  summarise(
    messageWordLength = sum(sapply(strsplit(text, " "), length))
  )
```

```{r}
  ggplot(chat_messages_word_count, aes(x = reorder(role, desc(role)), y = messageWordLength, fill = messageType)) + 
  geom_bar(stat = 'identity', position = 'stack') + 
  ggtitle("# of words produced by each player, according to utterance category") + 
  xlab('Role') + ylab('# of player\'s words')
```


## Informative Message Types
```{r}
informative_messages_count = chat_messages %>%
  filter(messageType == 'Informative') %>%
  select(utteranceType) %>%
  summarize(
    None = sum(utteranceType == 'None'),
    Quantifier = sum(utteranceType == 'Quantifier'),
    Conditional = sum(utteranceType == 'Conditional'),
    Exemplar = sum(utteranceType == 'Exemplar'),
    Generic = sum(utteranceType == 'Generic'),
    Imperative = sum(utteranceType == 'Imperative'),
    Affirmative = sum(utteranceType == 'Affirmative'),
    Knowledge = sum(utteranceType == 'Knowledge'),
    Logic = sum(utteranceType == 'Logic'),
    Question = sum(utteranceType == 'Question'),
    Possibility_Conditional = sum(utteranceType == 'Possibility,Conditional'),
  ) %>%
  select(None, Quantifier, Conditional, Exemplar, Generic, Imperative, Affirmative, Knowledge, Logic, Question, Possibility_Conditional) %>%
  gather(key = "stat", value = "val")
```

```{r fig.width=7, fig.height=4, echo=FALSE}
ggplot(
    informative_messages_count,
    aes(x = stat, y = val)
  ) + geom_bar(stat = 'identity') + ggtitle("# of Informative Messages, Across Utterance Type") +
    xlab('Utterance Type') + ylab('# of Messages Produced')
```

```{r}
informative_messages_word_count = chat_messages %>%
  filter(messageType == 'Informative') %>%
  select(utteranceType, text) %>%
  group_by(utteranceType) %>%
  summarise(
    messageWordLength = sum(sapply(strsplit(text, " "), length))
  )
```

```{r fig.width=7, fig.height=4, echo=FALSE}
ggplot(
    informative_messages_word_count,
    aes(x = utteranceType, y = messageWordLength)
  ) + geom_bar(stat = 'identity') + ggtitle("# of Informative Words, Across Utterance Type") +
    xlab('Utterance Type') + ylab('# of Words Produced')
```


## Follow-Up Message Types
```{r}
follow_up_messages_count = chat_messages %>%
  filter(messageType == 'Follow-Up') %>%
  mutate(role = as.character(role)) %>%
  mutate(role = replace(role, role == "explorer", "T")) %>%
  mutate(role = replace(role, role == "student", "S")) %>%
  group_by(role) %>%
  select(role, utteranceType) %>%
  summarize(
    None = sum(utteranceType == 'None'),
    Quantifier = sum(utteranceType == 'Quantifier'),
    Conditional = sum(utteranceType == 'Conditional'),
    Exemplar = sum(utteranceType == 'Exemplar'),
    Generic = sum(utteranceType == 'Generic'),
    Imperative = sum(utteranceType == 'Imperative'),
    Affirmative = sum(utteranceType == 'Affirmative'),
    Knowledge = sum(utteranceType == 'Knowledge'),
    Logic = sum(utteranceType == 'Logic'),
    Question = sum(utteranceType == 'Question'),
    Possibility_Conditional = sum(utteranceType == 'Possibility,Conditional'),
  ) %>%
  select(role, None, Quantifier, Conditional, Exemplar, Generic, Imperative, Affirmative, Knowledge, Logic, Question, Possibility_Conditional) %>%
  gather(key = "stat", value = "val", -role)
```


```{r fig.width=7, fig.height=4, echo=FALSE}
ggplot(
    follow_up_messages_count %>%
      select(val, stat) %>%
      group_by(stat) %>%
      summarize(val = sum(val)),
    aes(x = stat, y = val)
  ) + geom_bar(stat = 'identity') + ggtitle("# of Follow-Up Messages, Across Utterance Type") +
    xlab('Message Type') + ylab('# of Messages Produced')
```

```{r fig.width=7, fig.height=4, echo=FALSE}
ggplot(
    follow_up_messages_count,
    aes(x = reorder(role, desc(role)), y = val, fill=stat)
  ) + geom_bar(stat = 'identity', position = 'stack') + ggtitle("# of Follow-Up Messages, Across Utterance Type") +
    xlab('Utterance Type') + ylab('# of Messages Produced')
```

## Quantifier Specific Language
```{r}
quantifier_messages_count = chat_messages %>%
  filter(
    utteranceType == 'Quantifier' | 
      (utteranceType == 'Question' & 
         (grepl('any', text)  | 
            grepl('all', text) |
            grepl('mostly', text) | 
            grepl('some', text)
          )
       )
  ) %>%
  mutate(utteranceTypeSignifiers=replace(utteranceTypeSignifiers, utteranceType=='Question', 'Question')) %>%
  select(role, utteranceTypeSignifiers) %>%
  group_by(role, utteranceTypeSignifiers) %>%
  summarize(
    Count = length(utteranceTypeSignifiers)
  )
```

```{r fig.width=7, fig.height=4, echo=FALSE}
ggplot(
    quantifier_messages_count,
    aes(x = utteranceTypeSignifiers, y = Count)
  ) + geom_bar(stat = 'identity') + ggtitle("# of Messages with Quantifier Signifier (Teacher & Student Combined)") +
    xlab('Signifier') + ylab('# of Times Utilized')
```


```{r fig.width=7, fig.height=4, echo=FALSE}
ggplot(
    quantifier_messages_count,
    aes(x = role, y = Count, fill=utteranceTypeSignifiers)
  ) + geom_bar(stat = 'identity', position ='stack') + ggtitle("# of Messages with Quantifier Signifier (Teacher & Student Split)") +
    xlab('Role') + ylab('# of Times Utilized')
```

## Analysis of Langauge Across Pairs
```{r}
game_informative_msgs = chat_messages %>%
  filter(messageType == 'Informative') %>%
  group_by(gameid) %>%
  select(gameid, utteranceType) %>%
  summarize(
    None = sum(utteranceType == 'None'),
    Quantifier = sum(utteranceType == 'Quantifier'),
    Conditional = sum(utteranceType == 'Conditional'),
    Exemplar = sum(utteranceType == 'Exemplar'),
    Generic = sum(utteranceType == 'Generic'),
    Imperative = sum(utteranceType == 'Imperative'),
    Affirmative = sum(utteranceType == 'Affirmative'),
    Knowledge = sum(utteranceType == 'Knowledge'),
    Logic = sum(utteranceType == 'Logic'),
    Question = sum(utteranceType == 'Question'),
    Possibility_Conditional = sum(utteranceType == 'Possibility,Conditional'),
  ) %>%
  select(gameid, None, Quantifier, Conditional, Exemplar, Generic, Imperative, Affirmative, Knowledge, Logic, Question, Possibility_Conditional) %>%
  gather(key = "stat", value = "val", -gameid)
```

```{r fig.width=7, fig.height=4, echo=FALSE}
ggplot(
    game_informative_msgs,
    aes(x='', y=val, fill=stat)
  ) + geom_bar(stat = 'identity', position ='stack') + ggtitle("Distribution of Utterance Types (Informative) Across Teachers") +
    xlab('Teacher') + ylab('# of Messages of Utterance Type') + facet_grid(~gameid)
```

```{r}
pair_messages = chat_messages %>%
  filter(messageType == 'Follow-Up' | messageType == 'Informative') %>%
  mutate(role = as.character(role)) %>%
  mutate(role = replace(role, role == "explorer", "T")) %>%
  mutate(role = replace(role, role == "student", "S")) %>%
  group_by(gameid, role) %>%
  select(role, gameid, utteranceType) %>%
  summarize(
    None = sum(utteranceType == 'None'),
    Quantifier = sum(utteranceType == 'Quantifier'),
    Conditional = sum(utteranceType == 'Conditional'),
    Exemplar = sum(utteranceType == 'Exemplar'),
    Generic = sum(utteranceType == 'Generic'),
    Imperative = sum(utteranceType == 'Imperative'),
    Affirmative = sum(utteranceType == 'Affirmative'),
    Knowledge = sum(utteranceType == 'Knowledge'),
    Logic = sum(utteranceType == 'Logic'),
    Question = sum(utteranceType == 'Question'),
    Possibility_Conditional = sum(utteranceType == 'Possibility,Conditional'),
  ) %>%
  select(role, gameid, None, Quantifier, Conditional, Exemplar, Generic, Imperative, Affirmative, Knowledge, Logic, Question, Possibility_Conditional) %>%
  gather(key = "stat", value = "val", 3:13)
```

```{r fig.width=7, fig.height=4, echo=FALSE}
ggplot(
    pair_messages,
    aes(x=role, y=val, fill=stat)
  ) + geom_bar(stat = 'identity', position ='stack') + ggtitle("Distribution of Utterance Types (Informative + Follow-Up) Across Player Pairs ") +
    xlab('Role') + ylab('# of Messages of Utterance Type') + facet_grid(~gameid) + scale_color_viridis(discrete=TRUE)
```

## Analysis of Language Across Concepts

## Analysis of Language Across Concept Types 


