---
title: "MP-Game-5: Multiplayer Boolean Concept Learning (Summurization Format)"
output: beamer_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

# Data Processing
```{r}

library(reshape2)
library(purrr)
library(jsonlite)
library(tidyr) 
library(dplyr)
library(ggplot2)
library(scales)
library(rwebppl)
```

```{r}
#-----------------------
# Reading Mp-Game-3 Data
#-----------------------
# Load Training Trial Responses
temp <- list.files(
  "../../../data/mp-game-3/experiment/train_trials",
  pattern="*.csv",
  full.names=TRUE
)
train_trials_rbr <- do.call(rbind, lapply(temp, read.csv))

# Load Test Trial Responses
temp <- list.files(
  "../../../data/mp-game-3/experiment/test_trials",
  pattern="*.csv",
  full.names=TRUE
)
test_trials_rbr <- do.call(rbind, lapply(temp, read.csv))

# Load Training Summary Stats
temp <- list.files(
  "../../../data/mp-game-3/experiment/train_summary_stats",
  pattern="*.csv",
  full.names=TRUE
)
train_stats_rbr <- do.call(rbind, lapply(temp, read.csv))

# Load Test Summary Stats
temp <- list.files(
  "../../../data/mp-game-3/experiment/test_summary_stats",
  pattern="*.csv",
  full.names=TRUE
)
test_stats_rbr <- do.call(rbind, lapply(temp, read.csv))
```

```{r}
#-----------------------
# Reading Mp-Game-5 Data
#-----------------------
# Load Training Trial Responses
temp <- list.files(
  "../../../data/mp-game-5/logTrain",
  pattern="*.csv",
  full.names=TRUE
)
train_trials_s <- do.call(rbind, lapply(temp, read.csv, sep=""))

# Load Test Trial Responses
temp <- list.files(
  "../../../data/mp-game-5/logTest",
  pattern="*.csv",
  full.names=TRUE
)
test_trials_s <- do.call(rbind, lapply(temp, read.csv, sep=""))

# Load Test Summary Stats
temp <- list.files(
    "../../../data/mp-game-5/logScores",
  pattern="*.csv",
  full.names=TRUE
)
test_stats_s <- do.call(rbind, lapply(temp, read.csv, sep=""))
```


```{r}
rule_indices <- c(0, 2, 4, 6, 8)

#--------------------------------------
# Accuracy Calculations Round-By-Round
# -------------------------------------
test_acc_teacher_rbr <- test_stats_rbr %>%
  filter(role == "explorer") %>%
  mutate(correct = hits + correct_rejections) %>%
  mutate(incorrect = misses + false_alarms) %>%
  select(rule_idx, correct, incorrect) %>%
  filter(rule_idx %in% rule_indices) %>%
  group_by(rule_idx) %>% 
  summarise_all(sum) %>%
  mutate(acc = correct/(correct + incorrect)) %>%
  mutate(game = "round")
test_acc_teacher_rbr$rule_idx <- as.factor(test_acc_teacher_rbr$rule_idx)
test_acc_teacher_rbr$game <- as.factor(test_acc_teacher_rbr$game)

test_acc_listener_rbr <- test_stats_rbr %>%
  filter(role == "student") %>%
  mutate(correct = hits + correct_rejections) %>%
  mutate(incorrect = misses + false_alarms) %>%
  select(rule_idx, incorrect, correct) %>%
  filter(rule_idx %in% rule_indices) %>%
  group_by(rule_idx) %>% 
  summarise_all(sum) %>%
  mutate(acc = correct/(correct + incorrect)) %>%
  mutate(game = "round")
test_acc_listener_rbr$rule_idx <- as.factor(test_acc_listener_rbr$rule_idx)
test_acc_listener_rbr$game <- as.factor(test_acc_listener_rbr$game)

#--------------------------------------
# Accuracy Calculations Summurization
# -------------------------------------
test_acc_teacher_s <- test_stats_s %>%
  filter(role == "explorer") %>%
  mutate(correct = hits + correct_rejections) %>%
  mutate(incorrect = misses + false_alarms) %>%
  select(rule_idx, correct, incorrect) %>%
  filter(rule_idx %in% rule_indices) %>%
  group_by(rule_idx) %>% 
  summarise_all(sum) %>%
  mutate(acc = correct/(correct + incorrect)) %>%
  mutate(game = "summurization")
test_acc_teacher_s$game <- as.factor(test_acc_teacher_s$game)

test_acc_listener_s <- test_stats_s %>%
  filter(role == "student") %>%
  mutate(correct = hits + correct_rejections) %>%
  mutate(incorrect = misses + false_alarms) %>%
  select(rule_idx, incorrect, correct) %>%
  filter(rule_idx %in% rule_indices) %>%
  group_by(rule_idx) %>% 
  summarise_all(sum)%>%
  mutate(acc = correct/(correct + incorrect)) %>%
  mutate(game = "summurization")
test_acc_listener_s$game <- as.factor(test_acc_listener_s$game)

#----------------------------
# Accuracy Calculations Joint
# ---------------------------
test_acc_teacher_combined = bind_cols(
  test_acc_teacher_rbr %>%
    select(rule_idx, acc) %>%
    rename(round_by_round_acc = acc),
  test_acc_teacher_s %>%
    select(rule_idx, acc) %>%
    rename(summurization_acc = acc)
  ) %>%
  select(rule_idx, round_by_round_acc, summurization_acc)

test_acc_listener_combined = bind_cols(
  test_acc_listener_rbr %>%
    select(rule_idx, acc) %>%
    rename(round_by_round_acc = acc),
  test_acc_listener_s %>%
    select(rule_idx, acc) %>%
    rename(summurization_acc = acc)
  ) %>%
  select(rule_idx, round_by_round_acc, summurization_acc)
```

## Intro
* Goal: Collect cheaper data for cultural ratchet experiment.
* In spring, we ran a pilot experiment, with a "round-by-round" format, where players only learned a single concept.
* In fall, we ran a pilot experiment with a "summurization" format, where palyers played multiple concepts.
* For spring pilot (round-by-round), we had 5 concepts, each with two different lists of stimuli.
* For fall pilot (summurization), we had the same 5 concepts and selected one of two lists of stimuli, from spring.


## Cost Comparison
Spring Pilot Cost Per Round (Round-By-Round-Format):

  $122.40 / 36 = $3.40 / round
  
Fall Pilot Cost Per Round (Summurization-Format):

  ($(1.25 * 20 + .50 * 20) + $11.60)) / 40 = $1.165 / round

If we were maximally efficient, the Fall Pilot would be $.892 / round. We had two games crash midway because of players disconnecting; so our cost for the pilot was slightly higher than expected according to the data we collected.

## Experiment Details (Critters)
* 81 total possible critters: ~50 training, ~31 test
* 4 axes of variability:
    + Critter Type (Bug, Fish, Bird)
    + Primary Color (Blue, Green, Orange)
    + Secondary Color (Red, Yellow, Purple)
    + Size (Small, Medium, Large)


## Experiment Details (Concepts)
* Primary Color == Orange ("Orange things")
* Critter Type == Fish && Primary Color == Blue ("Blue fish")
* Primary Color == Orange && Secondary Color == Purple ("Purple and orange things")
* Critter Type == Bug || Secondary Color == Yellow ("Bugs, or yellow things")
* Critter Type == Bird || Primary Color == Green ("Birds, or green things")


## Explorer Performance (Round vs. Summurization Acc.)
```{r}
ggplot(test_acc_teacher_combined, aes(x=round_by_round_acc, y=summurization_acc), fill=rule_idx) +
  geom_point() +
  xlim(0, 1) +
  ylim(0, 1)
```


## Student Performance (Round vs. Summurization Acc.)


## Explorer Performance ()


## Next Steps
* Language Analysis
  
