---
title: "Critter Game - Data - Text Analysis"
author: "Lauren Oey"
date: "7/7/2017"
output: github_document
---

```{r preamble}
library(dplyr)
library(tidyr)
library(ggplot2)
library(data.table)
library(tidytext)

#setting the directory
setwd("../mturk/game-1/")


text_data <- read.csv("game-1-trials.csv")

reorder_size <- function(x) {
  factor(x, levels = names(sort(table(x), decreasing = TRUE)))
}

#View(text_data)

str(text_data)
summary(text_data)


# Adjusting MechTurk/Punctuation Errors
text_df <- data_frame(line = 1:18, condition=text_data[,4], text=text_data[,3])
text_df$text <- gsub("&quotechar", "'", text_df$text)
text_df$text <- gsub("DON'T.NONE", "DON'T. NONE", text_df$text)
text_df$text <- as.character(text_df$text)

```

# Tokenize Sentences

```{r}
#Tokenize text responses
tokenized <- text_df %>%
  unnest_tokens(word, text) # #convert sentence string to tokens
tokenized

#Create subset of tokenized data frame by condition
tokenized_detector <- subset(tokenized, text_df$condition=="pepsin_detector")
tokenized_book <- subset(tokenized, text_df$condition=="label_book")

#tokenized_byCond <- bind_rows(
  #mutate(tokenized_detector, condition="Pepsin Detector"),
  #mutate(tokenized_book, condition="Labelled Book"))

```

# All Words Count

```{r}
#Count of all words
count_all_words <- tokenized %>%
  count(word, sort=TRUE)
count_all_words

#Count of all words in Pepsin Detector condition
count_all_words_detector <- tokenized_detector %>%
  count(word, sort=TRUE)
count_all_words_detector

#Count of all words in Label Book condition
count_all_words_book <- tokenized_book %>%
  count(word, sort=TRUE)
count_all_words_book

count_all_words_byCond <- bind_rows(
  mutate(count_all_words_detector, condition="Pepsin Detector"),
  mutate(count_all_words_book, condition="Labelled Book"))
count_all_words_byCond
```


# Quantifiers Word Count

## Both Conditions Combined

```{r}
quantifier_words = c("all","each","most","many","some","few","none")
# no "every","a lot"

quantifiers <- count_all_words[c(count_all_words$word %in% quantifier_words),]
quantifiers

quantifier_count <- ggplot(quantifiers, aes(x=word, y=n)) +
  geom_bar(stat="identity", fill="black") +
  scale_x_discrete("Word Used in Response", limits=c(quantifier_words)) +
  scale_y_continuous("Frequency", limits=c(0,20)) + 
  ggtitle("Quantifiers") +
  theme_minimal() +
  theme(plot.title = element_text(hjust=0.5))

quantifier_count
```


## By Condition

```{r}
quantifiers_byCond <- count_all_words_byCond[c(count_all_words_byCond$word %in% quantifier_words),]
quantifiers_byCond


quantifier_count_byCond <- ggplot(quantifiers_byCond, aes(x=word, y=n)) +
  geom_bar(stat="identity", fill="black") +
  scale_x_discrete("Word Used in Response", limits=c(quantifier_words)) +
  scale_y_continuous("Frequency", limits=c(0,15)) + 
  facet_wrap(~condition) +
  ggtitle("Quantifiers") +
  theme_bw() +
  theme(plot.title = element_text(hjust=0.5, size=18, face="bold"),
        strip.background = element_rect(fill="skyblue"),
        strip.text = element_text(size=12, face="bold"))

quantifier_count_byCond
```


# Probabilistic Word Count

## Both Conditions Combined

```{r}
probability_words = c("always","usually","often","likely","sometimes","never")

probabilities <- count_all_words[c(count_all_words$word %in% probability_words),]
probabilities

probability_count <- ggplot(probabilities, aes(x=word, y=n)) +
  geom_bar(stat="identity", fill="black") +
  scale_x_discrete("Word Used in Response", limits=c(probability_words)) +
  scale_y_continuous("Frequency", limits=c(0,5)) + 
  ggtitle("Probability Words") +
  theme_minimal() +
  theme(plot.title = element_text(hjust=0.5))

probability_count
```

## By Condition

```{r}
probabilities_byCond <- count_all_words_byCond[c(count_all_words_byCond$word %in% probability_words),]
probabilities_byCond


probability_count_byCond <- ggplot(probabilities_byCond, aes(x=word, y=n)) +
  geom_bar(stat="identity", fill="black") +
  scale_x_discrete("Word Used in Response", limits=c(probability_words)) +
  scale_y_continuous("Frequency", limits=c(0,20)) + 
  facet_wrap(~condition) +
  ggtitle("Probability Words") +
  theme_bw() +
  theme(plot.title = element_text(hjust=0.5, size=18, face="bold"),
        strip.background = element_rect(fill="skyblue"),
        strip.text = element_text(size=12, face="bold"))

probability_count_byCond

# only present in 1 condition
```


# Conditionals Word Count

## Both Conditions Combined

```{r}
conditional_words = c("if","when","while")

conditionals <- count_all_words[c(count_all_words$word %in% conditional_words),]
conditionals

conditional_count <- ggplot(conditionals, aes(x=word, y=n)) +
  geom_bar(stat="identity", fill="black") +
  scale_x_discrete("Word Used in Response", limits=c(conditional_words)) +
  scale_y_continuous("Frequency", limits=c(0,10)) + 
  ggtitle("Conditionals") +
  theme_minimal() +
  theme(plot.title = element_text(hjust=0.5))

conditional_count

```

## By Condition

```{r}
conditionals_byCond <- count_all_words_byCond[c(count_all_words_byCond$word %in% conditional_words),]
conditionals_byCond


conditional_count_byCond <- ggplot(conditionals_byCond, aes(x=word, y=n)) +
  geom_bar(stat="identity", fill="black") +
  scale_x_discrete("Word Used in Response", limits=c(conditional_words)) +
  scale_y_continuous("Frequency", limits=c(0,20)) + 
  facet_wrap(~condition) +
  ggtitle("Conditionals") +
  theme_bw() +
  theme(plot.title = element_text(hjust=0.5, size=18, face="bold"),
        strip.background = element_rect(fill="skyblue"),
        strip.text = element_text(size=12, face="bold"))

conditional_count_byCond
```


# Context Word Count

## Both Conditions Combined

```{r}
count_context_words <- tokenized %>%
  anti_join(stop_words) %>% #remove function words
  count(word, sort=TRUE)
count_context_words

# first 15 words - no change to singular/plural
context_count_unfiltered <- ggplot(count_context_words[1:15,], aes(x=word, y=n)) +
  geom_bar(stat="identity", fill="black") +
  scale_x_discrete("Word Used in Response", limits=c(count_context_words$word[1:15])) +
  scale_y_continuous("Frequency", limits=c(0,18)) + 
  ggtitle("Context Words - Unfiltered") +
  theme_minimal() +
  theme(axis.text.x=element_text(angle=60, hjust=1),
        plot.title = element_text(hjust=0.5))

context_count_unfiltered
```

## By Condition

```{r}
#Count of all words in Pepsin Detector condition
count_context_words_detector <- tokenized_detector %>%
  anti_join(stop_words) %>% #remove function words
  count(word, sort=TRUE)
count_context_words_detector

#Count of all words in Label Book condition
count_context_words_book <- tokenized_book %>%
  anti_join(stop_words) %>% #remove function words
  count(word, sort=TRUE)
count_context_words_book

count_context_words_byCond <- bind_rows(
  mutate(count_context_words_detector, condition="Pepsin Detector"),
  mutate(count_context_words_book, condition="Labelled Book"))
count_context_words_byCond

# first 15 words - no change to singular/plural
context_count_unfiltered <- ggplot(count_context_words[1:15,], aes(x=word, y=n)) +
  geom_bar(stat="identity", fill="black") +
  scale_x_discrete("Word Used in Response", limits=c(count_context_words$word[1:15])) +
  scale_y_continuous("Frequency", limits=c(0,18)) + 
  ggtitle("Context Words - Unfiltered") +
  theme_minimal() +
  theme(axis.text.x=element_text(angle=60, hjust=1),
        plot.title = element_text(hjust=0.5))

context_count_unfiltered
```


# Convert Plural Nouns to Singular Nouns

```{r}
#changing plural to singular
text_noPlurals <- text_df
text_noPlurals$text <- gsub("feathers", "feather", text_noPlurals$text, ignore.case=TRUE)
text_noPlurals$text <- gsub("bones", "bone", text_noPlurals$text)
text_noPlurals$text <- gsub("creatures", "creature", text_noPlurals$text)
text_noPlurals$text <- gsub("critters", "critter", text_noPlurals$text)
text_noPlurals$text <- gsub("features", "feature", text_noPlurals$text)
text_noPlurals$text <- gsub("birds", "bird", text_noPlurals$text)
text_noPlurals$text <- gsub("colors", "color", text_noPlurals$text)
text_noPlurals$text <- gsub("tails", "tail", text_noPlurals$text, ignore.case=TRUE)
text_noPlurals$text <- gsub("heads", "head", text_noPlurals$text)
text_noPlurals$text <- gsub("wugs", "wug", text_noPlurals$text, ignore.case=TRUE)
text_noPlurals$text <- gsub("blickets", "blicket", text_noPlurals$text)
text_noPlurals$text <- gsub("rambos", "rambo", text_noPlurals$text, ignore.case=TRUE)
text_noPlurals$text <- gsub("peppin", "pepsin", text_noPlurals$text)
text_noPlurals$text <- gsub("pespin", "pepsin", text_noPlurals$text)
text_noPlurals$text <- gsub("pebcide", "pepsin", text_noPlurals$text)
text_noPlurals$text <- gsub("sizes", "size", text_noPlurals$text)
text_noPlurals$text <- gsub("types", "type", text_noPlurals$text)


tokenized_noPlurals <- text_noPlurals %>%
  unnest_tokens(word, text)

count_context_words_noPlurals <- tokenized_noPlurals %>%
  anti_join(stop_words) %>%
  count(word, sort=TRUE)
count_context_words_noPlurals


# first 15 words - w/ change to singular/plural
context_count_filtered <- ggplot(count_context_words_noPlurals[1:15,], aes(x=word, y=n)) +
  geom_bar(stat="identity", fill="black") +
  scale_x_discrete("Word Used in Response", limits=c(count_context_words_noPlurals$word[1:15])) +
  scale_y_continuous("Frequency", limits=c(0,18)) + 
  ggtitle("Context Words - Filtered") +
  theme_minimal() +
  theme(axis.text.x=element_text(angle=60, hjust=1),
        plot.title = element_text(hjust=0.5))

context_count_filtered
```

# Category Names

```{r}
# Categories = pepsin, tail, head, species, color, size
category_types = c("pepsin","tail","head","species","color","size")

characteristic_categories <- count_context_words_noPlurals[c(count_context_words_noPlurals$word %in% category_types),]
categories_count <- ggplot(characteristic_categories, aes(x=word, y=n)) +
  geom_bar(stat="identity", fill="black") +
  scale_x_discrete("Word Used in Response", limits=c(characteristic_categories$word[1:6])) +
  scale_y_continuous("Frequency", limits=c(0,18)) + 
  ggtitle("Critter Characteristic") +
  theme_minimal() +
  theme(plot.title = element_text(hjust=0.5))

categories_count
```

# Types Within Categories

## Species Categories

```{r}
# Species = rambo, wug, blicket
species_types = c("rambo","wug","blicket")
species_categories <- count_context_words_noPlurals[c(count_context_words_noPlurals$word %in% species_types),]
species_count <- ggplot(species_categories, aes(x=word, y=n)) +
  geom_bar(stat="identity", fill="black") +
  scale_x_discrete("Word Used in Response", limits=c(species_categories$word[1:3])) +
  scale_y_continuous("Frequency", limits=c(0,5)) + 
  ggtitle("Species Names") +
  theme_minimal() +
  theme(plot.title = element_text(hjust=0.5))

species_count
```

## Colors Categories

```{r}
# Species = rambo, wug, blicket
color_types = c("green","yellow","orange")
color_assignment = c("green" = "forestgreen", "yellow" = "gold", "orange" = "sienna1")
color_categories <- count_context_words_noPlurals[c(count_context_words_noPlurals$word %in% color_types),]
color_count <- ggplot(color_categories, aes(x=word, y=n, fill=word)) +
  geom_bar(stat="identity", colour="black") +
  scale_x_discrete("Word Used in Response", limits=c(color_categories$word[1:3])) +
  scale_y_continuous("Frequency", limits=c(0,5)) + 
  scale_fill_manual(values=color_assignment) +
  ggtitle("Color Descriptors") +
  guides(fill=FALSE) +
  theme_minimal() +
  theme(plot.title = element_text(hjust=0.5))

color_count
```



