---
title: "CritterGame_MP-1_TextAnalysis_Pilot_1"
author: "Lauren Oey, MH Tessler"
date: "8/13/2017 (updated 10/6/2017)" 
output: github_document
---

```{r preamble}
library(dplyr)
library(tidyr)
library(knitr)
library(ggplot2)
library(data.table)
library(tidytext)
#library(coreNLP)
#library(cleanNLP)
#library(reticulate)
library(xtable)
library(gridExtra)
library(forcats)
theme_set(theme_few())
```


Load message data

```{r}
df.chat <- rbindlist(
  lapply(list.files(path="../../data/mp-game-2/chatMessage/", 
                    pattern="*.csv", full.names=TRUE), fread), 
  fill=TRUE, idcol="dyad_id") %>%
  mutate(
    dyad_id = as.factor(ifelse(dyad_id < 10,
                            paste0("dyad_0", dyad_id),
                            paste0("dyad_", dyad_id))),
    gameid = as.factor(gameid),
    role = as.factor(role),
    trialNum = trialNum
  )

df.chat %>%
  select(dyad_id, trialNum, role, text) %>%
  kable()
```

Load training data

```{r}
df.train <- rbindlist(
  lapply(list.files(path="../../data/mp-game-2/logTrain/", 
                    pattern="*.csv", full.names=TRUE), fread), 
  fill=TRUE, idcol="dyad_id") %>%
  mutate(
    dyad_id = as.factor(ifelse(dyad_id < 10,
                            paste0("dyad_0", dyad_id),
                            paste0("dyad_", dyad_id))),
    gameid = as.factor(gameid),
    role = as.factor(role),
    trialNum = trialNum
  )

df.train.trials <- df.train %>%
  select(gameid,trialNum,  role,
         genus, conceptNumber, categoryPluralLabel, featureOrder, time_in_ms) %>%
  distinct()
```


Load and preprocess test data

```{r}
df.test <- rbindlist(lapply(list.files(path="../../data/mp-game-2/logTest/", pattern="*.csv", full.names=TRUE), fread), fill=TRUE, idcol="dyad_id") %>%
  mutate(
    dyad_id = as.factor(ifelse(dyad_id < 10,
                            paste0("dyad_0", dyad_id),
                            paste0("dyad_", dyad_id))),
    gameid = as.factor(gameid),
    role = as.factor(role),
    trialNum = trialNum - 1,
    correct = ifelse(categorizedResponse=="hit" |
                       categorizedResponse=="correctRejection", 1, 0),
    hit =  ifelse(categorizedResponse=="hit", 1, 0),
    falseAlarm = ifelse(categorizedResponse=="falseAlarm", 1, 0),
    correctRejection =  ifelse(categorizedResponse=="correctRejection", 1, 0),
    conceptNumber = factor(conceptNumber),
    featureOrder = factor(featureOrder),
    genus = factor(genus)
  )

df.test.trials <- df.test %>%
  group_by(gameid, trialNum, genus, categoryPluralLabel, conceptNumber, featureOrder, role,
           tested_on) %>%
  summarize(n_correct = sum(correct),
            prop_correct = sum(correct) / n(),
            n_hit = sum(hit),
            n_correctRejection = sum(correctRejection))
```



```{r}
df.chat.test <- left_join(
  left_join(
    df.chat %>% select(gameid, trialNum, role, text), 
    df.test.trials %>% ungroup() %>% select(-featureOrder) %>%#, -n_correct, -prop_correct) %>%
      rename(categoryTested = categoryPluralLabel,
             genusTested = genus,
             conceptTested = conceptNumber)
    ),
  df.train.trials %>%
    select(gameid, trialNum, role, 
           genus, conceptNumber, categoryPluralLabel) %>%
    rename(categoryTrained = categoryPluralLabel,
             genusTrained = genus,
             conceptTrained = conceptNumber)
) %>%
  mutate(dyad_id = factor(gameid, labels = paste('dyad_', 
                                                seq(0, length(unique(gameid))- 1),
                                                sep="")))
```


```{r}
df.trials <- df.chat.test %>%
  select(-text) %>%
  distinct()


  summarize(n_correct = sum(correct),
            prop_correct = sum(correct) / n(),
            n_hit = sum(hit),
            n_correctRejection = sum(correctRejection))
```



```{r}

df.chat.test.display <- df.chat.test %>% 
  select(dyad_id, trialNum,
         genusTrained, categoryTrained, conceptTrained, 
         role,
         text
         )

#df.chat.test.display[with(df.chat.test.display, order(conceptTrained)),] %>%
  df.chat.test.display %>%
  select(dyad_id, trialNum, role, conceptTrained, text) %>%
  kable()
```

```{r}
df.chat.test.input <- left_join(
  left_join(
    df.test.trials %>% 
      ungroup() %>% select(-featureOrder, 
                           -n_correct,
                           -prop_correct) %>%
      rename(categoryTested = categoryPluralLabel,
             genusTested = genus,
             conceptTested = conceptNumber),
    df.train.trials %>%
      select(dyad_id, trialNum, role, 
             genus, conceptNumber, categoryPluralLabel) %>%
      rename(categoryTrained = categoryPluralLabel,
               genusTrained = genus,
               conceptTrained = conceptNumber)
  ) %>% 
  mutate(language_input = 
           ifelse(tested_on == "self", 
                  ifelse(role == "a", "a", "b"),
                  ifelse(role == "a", "b", "a")
                  )
  ),
  df.chat %>% select(dyad_id, trialNum, role, text) %>%
    rename(language_input = role)
)
```

# Quantitative text analysis

Words by concept learned

```{r}
df.chat.test %>%
  unnest_tokens(word, text) %>%
  group_by(conceptTrained) %>%
  summarize(n_words = n()) %>%
  ggplot(., aes(x = conceptTrained, y = n_words))+
  geom_col()

ggsave(paste(fig.path, "words_concepts.pdf", sep= ""), height = 4, width = 9)
```

Words by dyad

```{r}
df.chat.test %>%
  filter(dyad_id %in% complete.participants$dyad_id) %>%
  unnest_tokens(word, text) %>%
  group_by(dyad_id) %>%
  summarize(n_words = n()) %>%
  ggplot(., aes(x = dyad_id, y = n_words))+
  geom_col()

ggsave(paste(fig.path, "words_dyads.pdf", sep= ""), height = 4, width = 9)
```

Removing stop words

```{r}
df.chat.test %>%
  unnest_tokens(word, text) %>%
  anti_join(stop_words) %>%
  group_by(conceptTrained) %>%
  summarize(n_words = n()) %>%
  ggplot(., aes(x = conceptTrained, y = n_words))+
  geom_col()
```


```{r echo=FALSE, results='hide', message=FALSE, warning=FALSE, tidy=TRUE}
# Tokenize Sentences

wordTypeDictionary <- list(
  quantifier = c("all","each","most","many","half","some","few","none"),
  numerics = c("1st","2nd","3rd","4th",
               "0","1","2","3","4",
               "first","second","third","fourth",
               "two","three","four",
               "50%","1/4","2/4","3/4","50/50","50","100"),
  probability = c("always","usually","often","likely","sometimes","rarely","never",
                  "chance","time","times"),
  conditionals = c("if","when","while"),
  #category_types = c("pepsin","tail","head","species","color","size")
  species = c("morseths","morseth",
              "ollers","oller",
              "kweps","kwep",
              "blins","blin",
              "reesles","reesle",
              "dorbs","dorb",
              "zorbs","zorb",
              "taifels","taifel",
              "trufts","truft",
              "daiths","daith",
              "mooks","mook",
              "frams","fram",
              "moxes","mox",
              "luzaks","luzak",
              "javs","jav",
              "pangolins","pangolin",
              "ackles","ackle",
              "wugs","wug",
              "cheebas","cheeba",
              "elleps","ellep",
              "kazzes","kaz",
              "lorches","lorch",
              "plovs","plov",
              "grinks","grink",
              "glippets","glippet",
              "sapers","saper",
              "stups","stup",
              "krivels","krivel",
              "zoovs","zoov",
              "thups","thup",
              "crullets","crullet",
              "feps","fep"),
  color = c("red","orange","yellow","green","blue","purple"),
  internal_prop = c("eggs","poisonous","crocodiles","leaves",
                    "egg","crocodile","leaf","crocs","clovers","clover",
                    "alligators","alligator","spiders","spider","poison",
                    "lay","lays","grow","grows","live","lives"),
  creatures = c("birds","bird","bugs","bug","trees","tree","fish",
                "animals","animal","plants","plant",
                "creatures","creature","critters","critter","types","type"),
  features = c("properties","property","attributes","attribute",
               "colors","color"),
  hedge=c("guess","think","sense"),
  conversation=c("ok","okay","kk","k","sure","u bet",
                 "ready","rdy","sorry","gg","hi",
                 "good luck","luck","here we go","cya","continue",
                 "thanx","thanks","ty",
                 "yup","got it","gotcha","cool","yep","yeah","yes","good",
                 "sounds like a plan","plan",
                 "lmk","make sense","tricky","tough","whenever","oh","wow","agree"),
  boolean_properties = c("whiskers", "tail","wings","spikes",
                        "legs","eyes","feathers", "head","tails",
                        "thorns"),
category_genera = c("bug","bird", "fish", "flower"),
adjectives = c("short", "wide","tall","fat","skinny")
)

wordTypeDict <- utils::stack(wordTypeDictionary)
f <- function(x) as.character(wordTypeDict[match(tolower(x), wordTypeDict[[1]]), 2])

genericDict <- list(
  quantifier = c("all","each","most","many","half","some","few","none"),
  numerics = c("50%","1/4","2/4","3/4","50/50",
               "1st","2nd","3rd","4th",
               "1","2","3","4",
               "first","second","third","fourth","last",
               "one","two","three","four"),
  probability = c("always","usually","often","likely","sometimes","never"),
  demonstrative = c("these", "this", "those") #excluding that because of complications
  )
g <- function(x) ifelse(token_coreNLP$lemma %in% genericDict[[1]], FALSE,
       ifelse(grepl("wug", token_coreNLP$lemma), FALSE, TRUE))

#Tokenize text responses
# tokenized_wCond <- df.chat %>%
#   unnest_tokens(word, text)


# tokenized_ungrouped <- text_df %>%
#   unnest_tokens(word, text) %>%
#   count(word, sort=TRUE)

#this seems to be broken :/
# tokenized_coreNLP <- token_coreNLP %>%
#   filter(upos != "PUNCT") %>%
#   right_join(tokenized_wCond, by="word") %>%
#   distinct(word, lemma, tid, upos, roundNum, sender)
  
  
# tokenized <- token_coreNLP %>%
#   count(lemma, sort=TRUE) %>%
#   rowwise() %>%
#   mutate(wordType = f(lemma))
# 
# tokenized_word <- token_coreNLP %>%
#   count(word, sort=TRUE) %>%
#   rowwise() %>%
#   mutate(wordType = f(word))
# 
# 
# tokenized_generic <- token_coreNLP %>%
#   mutate(genericType = ifelse(upos=="NUM", FALSE, g(lemma)),
#          sentence_id = gsub(" ", "", paste(as.character(id), "_", as.character(sid))))
#View(tokenized_generic)



my_stop_words <- stop_words %>%
  filter(!(
    (word %in% c(wordTypeDictionary$probability,
                 wordTypeDictionary$quantifier,
                 wordTypeDictionary$conditionals)
    )))

context_by_words_type <-  df.chat.test %>%
  unnest_tokens(word, text) %>%
  anti_join(my_stop_words) %>% #remove function words
  #group_by(conceptTrained) %>%
  count(word, sort=TRUE) %>%
  rowwise() %>%
  mutate(wordType = f(word),
         wordType = factor(wordType))
context_by_words_type

```
```{r}
df.chat.test %>%
  unnest_tokens(word, text) %>%
  filter(word %in% 
           c(wordTypeDictionary$quantifier,
             wordTypeDictionary$probability)) %>% 
  #group_by(conceptTrained) %>%
  count(word, sort=TRUE) 
```

# Tables

## Messages by Distribution & Game

```{r echo=FALSE, warning=FALSE}

text_byPartnerScore <- score_byText %>%
  select(distribution, gameid_num, trialNum, role, messages, partnerPercentCorrect, partnerPercentCorrectlyHit, partnerPercentCorrectlyRejected) %>%
  mutate(distribution = factor(
    distribution, levels = c("[0, 1, 0]","[0, 1, 0.75]","[0, 1, 0.25]","[0.5, 0.5, 0.5]"))) %>%
  arrange(distribution, gameid_num, trialNum, role, desc(partnerPercentCorrect)) %>%
  rename(dyad = gameid_num, trial = trialNum, player = role, totalCorrect = partnerPercentCorrect, hit = partnerPercentCorrectlyHit, CR = partnerPercentCorrectlyRejected) 
kable(text_byPartnerScore, format.args = list(digits=2))
```

## Messages by Distribution & Partner Accuracy

```{r echo=FALSE, warning=FALSE}
# Binning by Accuracy on Test & Count Words in Text
split_byDistr_byAccuracy <- score_byText %>%
  group_by(distribution) %>%
  mutate(accuracyBin = ifelse(partnerPercentCorrect >= median(partnerPercentCorrect), "upper_bin", "lower_bin")) %>%
  ungroup()

wordCount_byDistr_byAccuracy <- split_byDistr_byAccuracy %>%
  group_by(distribution, accuracyBin) %>%
  unnest_tokens(word, messages) %>%
  count(word, sort=TRUE) %>%
  summarise(totalWords = sum(n))
#wordCount_byDistr_byAccuracy


table_byAccuracyBin <- split_byDistr_byAccuracy %>%
  select(distribution, accuracyBin, gameid_num, trialNum, role, messages, partnerPercentCorrect, partnerPercentCorrectlyHit, partnerPercentCorrectlyRejected) %>%
  mutate(distribution = factor(
    distribution, levels = c("[0, 1, 0]","[0, 1, 0.75]","[0, 1, 0.25]","[0.5, 0.5, 0.5]"))) %>%
  arrange(distribution, desc(partnerPercentCorrect), gameid_num, trialNum, role) %>%
  rename(dyad = gameid_num, trial = trialNum, player = role, totalCorrect = partnerPercentCorrect, hit = partnerPercentCorrectlyHit, CR = partnerPercentCorrectlyRejected) 
kable(table_byAccuracyBin, format.args = list(digits=2))
```

# Word Count by Distribution & Partner Test Accuracy

```{r echo=FALSE, results='hide', message=FALSE, warning=FALSE, tidy=TRUE}
wordCount_byInput <- split_byDistr_byAccuracy %>%
  group_by(gameid_num, trialNum, role, distribution, accuracyBin) %>%
  unnest_tokens(word, messages) %>%
  count(word, sort=TRUE) %>%
  mutate(words_in_message = sum(n))
#wordCount_byInput

wordCount_byDistr_byAccuracy <- wordCount_byInput %>%
  group_by(distribution, accuracyBin) %>%
  filter(!is.na(distribution)) %>%
  summarise(avgWords = mean(words_in_message))
#wordCount_byDistr_byAccuracy

plot_wordCount <- ggplot(wordCount_byDistr_byAccuracy, aes(x=accuracyBin, y=avgWords)) +
  geom_bar(stat="identity", fill="black") +
  scale_x_discrete("Partner Test Accuracy", labels=c("< Median",">= Median")) +
  scale_y_continuous("Avg Words Per Round", expand=c(0,0)) +
  facet_wrap(~distribution) +
  ggtitle("Word Count in Round by Distribution & Accuracy") +
  #coord_flip() +
  theme_classic() + 
  theme(plot.title = element_text(hjust=0.5, size=18, face="bold"),
        axis.title = element_text(size=24),
        axis.text.x = element_text(size=14, face="bold"),
        axis.text.y = element_text(size=20),
        strip.text.x = element_text(size = 16))
plot_wordCount
#ggsave("~/Desktop/Research/Summer2017/CSLI/FinalPresentation/Graphics/avgWordCount.png", plot_wordCount, height=8, width=12)
```


# All Words Counted by Word Type (Not Excluding Function Words)

```{r echo=FALSE, warning=FALSE}
all_words_byType <- tokenized %>%
  group_by(wordType) %>%
  summarize(n = sum(n)) %>%
  ungroup() %>%
  mutate(wordType = ifelse(is.na(wordType), "other", wordType),
         wordType  = factor(wordType, levels = wordType[order(n)]) ) %>%
  ggplot(., aes( x = wordType, y = n))+
  geom_col(fill="black")+
  ggtitle("All Words by Type")+
  theme_minimal()+
  theme(axis.text.x = element_text(angle = 90),
        plot.title = element_text(hjust=0.5))
all_words_byType

# Word Count by Condition by Type
# tokenized %>%
#   group_by(condition, wordType) %>%
#   summarize(n = sum(n)) %>%
#   ungroup() %>%
#   mutate(wordType = ifelse(is.na(wordType), "other", wordType),
#          wordType  = factor(wordType, levels = wordType[order(n)]) ) %>%
#   ggplot(., aes( x = wordType, y = n ))+
#   geom_col()+
#   facet_wrap(~condition, scales = 'free')+
#   theme(axis.text.x = element_text(angle = 0))+
#   coord_flip()

```

```{r echo=FALSE, results='hide', message=FALSE, warning=FALSE, tidy=TRUE}
count_all_words_byDistr <-  text_df %>%
  unnest_tokens(word, text) %>%
  group_by(distribution) %>%
  count(word, sort=TRUE) %>%
  rowwise() %>%
  filter(n > 6) %>%
  mutate(wordType = f(word),
         word = factor(word, levels= word[order(-n)]))

all_mostFreq_byDistr <- ggplot(count_all_words_byDistr, aes(x=reorder(word, n), y=n)) +
  geom_bar(stat="identity", fill="black") +
  scale_x_discrete("Word Used in Response") +
  scale_y_continuous("Counts", expand=c(0,0)) +
  facet_wrap(~distribution) +
  ggtitle("Common Words by Distribution") +
  coord_flip() + 
  theme_classic() + 
  theme(plot.title = element_text(hjust=0.5, size=28, face="bold"),
        axis.title = element_text(size=24),
        axis.text.x = element_text(size=16, face="bold"),
        axis.text.y = element_text(size=12),
        strip.text.x = element_text(size = 16))
all_mostFreq_byDistr
```

# Context Words Only

```{r echo=FALSE, results='hide', message=FALSE, warning=FALSE, tidy=TRUE}
count_context_words <-  text_df %>%
  unnest_tokens(word, text) %>%
  anti_join(stop_words) %>% #remove function words
  count(word, sort=TRUE) %>%
  rowwise() %>%
  filter(n > 4) %>%
  mutate(wordType = f(word),
         word = factor(word, levels= word[order(-n)]))

mostFreq <- ggplot(count_context_words, aes(x=reorder(word, n), y=n)) +
  geom_bar(stat="identity", fill="black") +
  scale_x_discrete("Word Used in Response") +
  scale_y_continuous("Counts", expand=c(0,0)) +
  ggtitle("Most Common Context Words") +
  coord_flip() + 
  theme_classic() + 
  theme(plot.title = element_text(hjust=0.5, size=20, face="bold"),
        axis.title = element_text(size=24),
        axis.text.x = element_text(size=20, face="bold"),
        axis.text.y = element_text(size=20))
mostFreq

#ggsave("~/Desktop/Research/Summer2017/CSLI/FinalPresentation/Graphics/mostFrequent_context.png", mostFreq, height=8, width=12)
```


```{r echo=FALSE, results='hide', message=FALSE, warning=FALSE, tidy=TRUE}
count_context_words_byDistr <-  text_df %>%
  unnest_tokens(word, text) %>%
  anti_join(stop_words) %>% #remove function words
  group_by(distribution) %>%
  count(word, sort=TRUE) %>%
  rowwise() %>%
  filter(n > 3) %>%
  mutate(wordType = f(word),
         word = factor(word, levels= word[order(-n)]))

context_mostFreq_byDistr <- ggplot(count_context_words_byDistr, aes(x=reorder(word, n), y=n)) +
  geom_bar(stat="identity", fill="black") +
  scale_x_discrete("Word Used in Response") +
  scale_y_continuous("Counts", expand=c(0,0), limits=c(0,12)) +
  facet_wrap(~distribution) +
  ggtitle("Common Context Words by Distribution") +
  coord_flip() + 
  theme_classic() + 
  theme(plot.title = element_text(hjust=0.5, size=14, face="bold"),
        axis.title = element_text(size=24),
        axis.text.x = element_text(size=12),
        axis.text.y = element_text(size=16),
        strip.text.x = element_text(size = 16))
context_mostFreq_byDistr
```

```{r echo=FALSE, results='hide', message=FALSE, warning=FALSE, tidy=TRUE}
count_context_words_type <-  text_df %>%
  unnest_tokens(word, text) %>%
  anti_join(stop_words) %>% #remove function words
  count(word, sort=TRUE) %>%
  rowwise() %>%
  mutate(wordType = f(word),
         wordType = factor(wordType)) %>%
  ungroup() %>%
  group_by(wordType) %>%
  summarize(total = sum(n)) %>%
  ungroup() %>%
  mutate(wordType = factor(wordType, levels= wordType[order(-total)]))

wordTypeKey <- c(
  quantifier = "quantifier",
  numerics = "numerics",
  probability = "probability",
  conditionals = "conditionals",
  species = "species",
  color = "color",
  internal_prop = "internal property",
  creatures = "creature (general)",
  features = "feature (general)",
  hedge="hedge",
  conversation="conversation")

context_byWordtype <- ggplot(count_context_words_type, aes(x=wordType, y=total)) +
  geom_bar(stat="identity", fill="black") +
  scale_x_discrete("Word Types", labels=wordTypeKey) +
  scale_y_continuous("Counts", expand=c(0,0)) +
  ggtitle("Context Word Types") +
  theme_classic() + 
  theme(plot.title = element_text(hjust=0.5, size=36, face="bold"),
        axis.title = element_text(size=24),
        axis.text.x = element_text(size=18, angle=90),
        axis.text.y = element_text(size=20))
context_byWordtype
#ggsave("~/Desktop/Research/Summer2017/CSLI/FinalPresentation/Graphics/wordType_context.png", context_byWordtype, height=8, width=12)

```

```{r echo=FALSE, results='hide', message=FALSE, warning=FALSE, tidy=TRUE}
count_context_words_type_byDistr <-  text_df %>%
  unnest_tokens(word, text) %>%
  anti_join(stop_words) %>% #remove function words
  group_by(distribution) %>%
  count(word, sort=TRUE) %>%
  rowwise() %>%
  mutate(wordType = f(word),
         wordType = factor(wordType)) %>%
  ungroup() %>%
  group_by(wordType, distribution) %>%
  summarize(total = sum(n)) %>%
  ungroup() %>%
  mutate(wordType = factor(wordType, levels= wordType[order(-total)]))

context_byWordtype_byDistr <- ggplot(count_context_words_type_byDistr, aes(x=wordType, y=total)) +
  geom_bar(stat="identity", fill="black") +
  scale_x_discrete("Word Types") +
  scale_y_continuous("Counts", expand=c(0,0)) +
  facet_wrap(~distribution) +
  ggtitle("Context Word Types by Distribution") +
  theme_classic() + 
  theme(plot.title = element_text(hjust=0.5, size=36, face="bold"),
        axis.title = element_text(size=24),
        axis.text.x = element_text(size=18, angle=90),
        axis.text.y = element_text(size=20),
        strip.text.x = element_text(size = 16))
context_byWordtype_byDistr
```

# Quantifiers Word Count

## Distribution Conditions Combined

```{r echo=FALSE, results='hide', message=FALSE, warning=FALSE, tidy=TRUE}
quantifier_words = c("all","each","most","many","half","some","few","none")

quantifiers <- tokenized %>%
  filter(lemma %in% quantifier_words) %>%
  group_by(lemma) %>%
  summarize(n = sum(n))

quantifier_count <- ggplot(quantifiers, aes(x=lemma, y=n)) +
  geom_bar(stat="identity", fill="black") +
  scale_x_discrete("Word Used in Response", limits=c(quantifier_words)) +
  scale_y_continuous("Frequency", limits=c(0,20)) + 
  ggtitle("Quantifiers") +
  theme_minimal() +
  theme(plot.title = element_text(hjust=0.5))

quantifier_count
```

## By Distribution

```{r echo=FALSE, results='hide', message=FALSE, warning=FALSE, tidy=TRUE}
count_quantifiers_byDistr <-  text_df %>%
  unnest_tokens(word, text) %>%
  group_by(distribution) %>%
  count(word, sort=TRUE) %>%
  filter(word %in% quantifier_words) %>%
  group_by(distribution, word) %>%
  summarize(n = sum(n))

quantifiers_byDistr <- ggplot(count_quantifiers_byDistr, aes(x=word, y=n)) +
  geom_bar(stat="identity", fill="black") +
  scale_x_discrete("Word") +
  scale_y_continuous("Counts", expand=c(0,0)) +
  facet_wrap(~distribution) +
  ggtitle("Quantifiers by Distribution") +
  theme_classic() + 
  theme(plot.title = element_text(hjust=0.5, size=36, face="bold"),
        axis.title = element_text(size=24),
        axis.text.x = element_text(size=18, angle=90),
        axis.text.y = element_text(size=20),
        strip.text.x = element_text(size = 16))
quantifiers_byDistr
```

# Probabilistic Word Count

## Distribution Conditions Combined

```{r echo=FALSE, results='hide', message=FALSE, warning=FALSE, tidy=TRUE}
probability_words = c("always","usually","often","likely","sometimes","rarely","never")

probabilities <- tokenized %>%
  filter(lemma %in% probability_words) %>%
  group_by(lemma) %>%
  summarize(n = sum(n))

probability_count <- ggplot(probabilities, aes(x=lemma, y=n)) +
  geom_bar(stat="identity", fill="black") +
  scale_x_discrete("Word Used in Response", limits=c(probability_words)) +
  scale_y_continuous("Frequency") + 
  ggtitle("Probability Words") +
  theme_minimal() +
  theme(plot.title = element_text(hjust=0.5))

probability_count
```

# Conditionals Word Count

## Both Conditions Combined

```{r eecho=FALSE, results='hide', message=FALSE, warning=FALSE, tidy=TRUE}
conditional_words = c("if","when","while")

conditionals <- tokenized %>%
  filter(lemma %in% conditional_words) %>%
  group_by(lemma) %>%
  summarize(n = sum(n))

conditional_count <- ggplot(conditionals, aes(x=lemma, y=n)) +
  geom_bar(stat="identity", fill="black") +
  scale_x_discrete("Word Used in Response", limits=c(conditional_words)) +
  scale_y_continuous("Frequency", limits=c(0,2)) + 
  ggtitle("Conditionals") +
  theme_minimal() +
  theme(plot.title = element_text(hjust=0.5))

conditional_count

```

# Numeric Word Count

#### Collapsed distributions into synonymous distributions (e.g. 1/2 = 50/50 = 50%)

```{r echo=FALSE, results='hide', message=FALSE, warning=FALSE, tidy=TRUE}
numeric_words = c("50%","1/4","2/4","3/4","50/50",
             "1st","2nd","3rd","4th",
             "1","2","3","4",
             "first","second","third","fourth",
             "two","three","four")
numeric_word_order = c("1","2","3","4",
                       "1st","2nd","3rd","4th",
                       "1/4","1/2","3/4")

numerics <- tokenized %>%
  filter(lemma %in% numeric_words)
#numerics$lemma <- gsub("one", "1", numerics$lemma)
numerics$lemma <- gsub("two", "2", numerics$lemma)
numerics$lemma <- gsub("three", "3", numerics$lemma)
numerics$lemma <- gsub("four", "4", numerics$lemma)
numerics$lemma <- gsub("first", "1st", numerics$lemma)
numerics$lemma <- gsub("second", "2nd", numerics$lemma)
numerics$lemma <- gsub("third", "3rd", numerics$lemma)
numerics$lemma <- gsub("fourth", "4th", numerics$lemma)
numerics$lemma <- gsub("50/50", "1/2", numerics$lemma)
numerics$lemma <- gsub("50%", "1/2", numerics$lemma)
numerics$lemma <- gsub("2/4", "1/2", numerics$lemma)
numerics <- numerics %>%
  group_by(lemma) %>%
  summarize(n = sum(n))

numeric_count <- ggplot(numerics, aes(x=lemma, y=n)) +
  geom_bar(stat="identity", fill="black") +
  scale_x_discrete("Word Used in Response", limits=numeric_word_order) +
  scale_y_continuous("Frequency") + 
  ggtitle("Number Word Counts") +
  theme_minimal() +
  theme(plot.title = element_text(hjust=0.5))

numeric_count

```


# Types Within Categories

## Species Categories

```{r echo=FALSE, results='hide', message=FALSE, warning=FALSE, tidy=TRUE}
species_types = c("morseth","oller","kewp","blin","reesle","dorb","zorb",
              "taifel","truft","daith","mook","fram","mox","luzak","jav",
              "pangolin","ackle","wug","cheeba","ellep","kaz","lorch",
              "plov","grink","glippet","saper","stup","krivel","zoov",
              "thup","crullet","fep")

species_categories <- tokenized %>%
  filter(lemma %in% species_types) %>%
  group_by(lemma) %>%
  summarize(n = sum(n))

species_count <- ggplot(species_categories, aes(x=lemma, y=n)) +
  geom_bar(stat="identity", fill="black") +
  scale_x_discrete("Word Used in Response", limits=c(species_types)) +
  scale_y_continuous("Frequency") + 
  ggtitle("Species Names") +
  theme_minimal() +
  theme(plot.title = element_text(hjust=0.5),
        axis.text.x = element_text(angle = 90))

species_count
```

## Colors Categories

```{r echo=FALSE, results='hide', message=FALSE, warning=FALSE, tidy=TRUE}
color_types = c("red", "orange","yellow","green","blue","purple")

color_assignment = c("red" = "red", "orange" = "sienna1", "yellow" = "gold", "green" = "forestgreen", "blue" = "blue", "purple" = "purple")
#color_categories <- tokenized[c(tokenized$word %in% color_types),]

color_categories <- tokenized %>%
  filter(lemma %in% color_types) %>%
  group_by(lemma) %>%
  summarize(n = sum(n))

color_count <- ggplot(color_categories, aes(x=lemma, y=n, fill=lemma)) +
  geom_bar(stat="identity", colour="black") +
  scale_x_discrete("Word Used in Response", limits=c(color_types)) +
  scale_y_continuous("Frequency") + 
  scale_fill_manual(values=color_assignment) +
  ggtitle("Color Descriptors") +
  guides(fill=FALSE) +
  theme_minimal() +
  theme(plot.title = element_text(hjust=0.5))

color_count
```

## Internal Property Categories

```{r echo=FALSE, results='hide', message=FALSE, warning=FALSE, tidy=TRUE}

internalProp_types = c("egg","poisonous","crocodile","leaf","crocs","clover","alligator","poison")
internalProp_types2 = c("egg","poisonous","crocodile","leaf")
internalProp_categories <- tokenized %>%
  filter(lemma %in% internalProp_types)

internalProp_categories$lemma <- gsub("crocs", "crocodile", internalProp_categories$lemma)
internalProp_categories$lemma <- gsub("alligator", "crocodile", internalProp_categories$lemma)
internalProp_categories$lemma <- gsub("clover", "leaf", internalProp_categories$lemma)
internalProp_categories$lemma <- gsub("poisonous", "poison", internalProp_categories$lemma)
internalProp_categories$lemma <- gsub("poison", "poisonous", internalProp_categories$lemma)

internalProp_categories <- internalProp_categories %>%
  ungroup() %>%
  group_by(lemma) %>%
  summarize(n = sum(n))

internalProp_count <- ggplot(internalProp_categories, aes(x=lemma, y=n, fill=lemma)) +
  geom_bar(stat="identity", colour="black") +
  scale_x_discrete("Word Used in Response") +
  scale_y_continuous("Frequency") + 
  ggtitle("Internal Property Descriptors") +
  guides(fill=FALSE) +
  theme_minimal() +
  theme(plot.title = element_text(hjust=0.5))

internalProp_count
```


# Total Word Count by Speaker
```{r echo=FALSE, results='hide', message=FALSE, warning=FALSE, tidy=TRUE}
# by Speaker in a given game
all_words_byRole <- tokenized_wCond %>%
  count(role, gameid_num, sort=TRUE)

byRole_count <- ggplot(all_words_byRole, aes(x=role, y=n)) +
  geom_bar(stat="identity", fill="black") +
  scale_x_discrete("Role") +
  scale_y_continuous("Frequency") + 
  facet_wrap(~gameid_num) +
  ggtitle("Player Word Count") +
  guides(fill=FALSE) +
  theme_minimal() +
  theme(plot.title = element_text(hjust=0.5),
        strip.text.x = element_text(size = 16))

byRole_count
```

# Total Word Count by Trial
```{r echo=FALSE, results='hide', message=FALSE, warning=FALSE, tidy=TRUE}
# by Round in a given game
all_words_byTrial <- tokenized_wCond %>%
  count(trialNum, sort=TRUE)

byTrial_count <- ggplot(all_words_byTrial, aes(x=trialNum, y=n)) +
  geom_bar(stat="identity", fill="black") +
  scale_x_discrete("Trial Number", limits=c(1:4)) +
  scale_y_continuous("Frequency") + 
  ggtitle("Word Count by Trial Number") +
  guides(fill=FALSE) +
  theme_minimal() +
  theme(plot.title = element_text(hjust=0.5))

byTrial_count

```

# Total Word Count by POS
```{r echo=FALSE, results='hide', message=FALSE, warning=FALSE, tidy=TRUE}
# by POS
all_words_byPOS <- token_coreNLP %>%
  count(upos, sort=TRUE)

byPOS_count <- ggplot(all_words_byPOS, aes(x=reorder(upos, -n), y=n)) +
  geom_bar(stat="identity", fill="black") +
  scale_x_discrete("POS") +
  scale_y_continuous("Frequency") + 
  ggtitle("Word Count by POS") +
  guides(fill=FALSE) +
  theme_minimal() +
  theme(plot.title = element_text(hjust=0.5),
        axis.text.x = element_text(angle = 45))

byPOS_count

```

# Total Word Count by Dependencies
```{r echo=FALSE, results='hide', message=FALSE, warning=FALSE, tidy=TRUE}
# by Dependencies
all_words_byDep <- dependency_coreNLP %>%
  count(relation, sort=TRUE)

byDep_count <- ggplot(all_words_byDep, aes(x=reorder(relation, -n), y=n)) +
  geom_bar(stat="identity", fill="black") +
  scale_x_discrete("Dependency") +
  scale_y_continuous("Frequency") + 
  ggtitle("Word Count by Dependency") +
  guides(fill=FALSE) +
  theme_minimal() +
  theme(plot.title = element_text(hjust=0.5),
        axis.text.x = element_text(angle = 90))

byDep_count

```