---
title: "Cultural Ratchet Data Exploration"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(reshape2)
library(purrr)
library(jsonlite)
library(tidyr) 
library(dplyr)
library(tidyboot)
library(ggplot2)
library(scales)
library(stringr)
library(digest)
```

```{r}
#-----------------------
# Reading Mp-Game-6 Data
#-----------------------

# All Games

# Load Test Trial Responses
temp <- list.files(
  "../../../data/mp-game-6/all_games/logTest",
  pattern="*.csv",
  full.names=TRUE
)
test_all_games <- do.call(rbind, lapply(temp, read.csv, sep=""))

# Load Test Summary Stats
temp <- list.files(
    "../../../data/mp-game-6/all_games/logScores",
  pattern="*.csv",
  full.names=TRUE
)
scores_all_games <- do.call(rbind, lapply(temp, read.csv, sep=""))

# Load Chat Messages
temp <- list.files(
    "../../../data/mp-game-6/all_games/chatMessage",
  pattern="*.csv",
  full.names=TRUE
)
msgs_complete_games <-do.call(rbind, lapply(temp, readr::read_tsv)) %>%
  group_by(gameid)

# Complete Games

# Load Test Trial Responses
temp <- list.files(
  "../../../data/mp-game-6/complete_games/logTest",
  pattern="*.csv",
  full.names=TRUE
)
test_complete_games <- do.call(rbind, lapply(temp, read.csv, sep=""))

# Load Test Summary Stats
temp <- list.files(
    "../../../data/mp-game-6/complete_games/logScores",
  pattern="*.csv",
  full.names=TRUE
)
scores_complete_games <- do.call(rbind, lapply(temp, read.csv, sep=""))

# Load Chat Messages
temp <- list.files(
    "../../../data/mp-game-6/complete_games/chatMessage",
  pattern="*.csv",
  full.names=TRUE
)
msgs_complete_games <-do.call(rbind, lapply(temp, readr::read_tsv)) %>%
  group_by(gameid)
```

# Distribution of Complete Data
```{r}
num_complete_games_played <- scores_complete_games %>%
  count(rule_idx) %>%
  mutate(n = n/2)
num_complete_games_played$n <- as.factor(num_complete_games_played$n)
ggplot(data=num_complete_games_played, aes(x=rule_idx, y=n)) +
  geom_bar(stat="identity") +
  labs(
    title = "Data Distribution of Complete Games",
    x = "Rule Index",
    y = "Number of Samples"
  )  +
  ylim(0, 10) +
  scale_y_discrete()
```

# Distribution of All Data
```{r}
num_all_games_played <- scores_all_games %>%
  count(rule_idx) %>%
  mutate(n = ceiling(n/2))
num_all_games_played$n <- as.factor(num_all_games_played$n)
ggplot(data=num_all_games_played, aes(x=rule_idx, y=n), ) +
  geom_bar(stat="identity") +
  labs(
    title = "Data Distribution of All Games",
    x = "Rule Index",
    y = "Number of Samples"
  ) +
  scale_y_discrete()
```

```

# Student Performance Vs. Teacher Performance (Complete Game)
```{r}
# Accuracies
test_explorer <- scores_complete_games %>%
  filter(role == "explorer") %>%
  mutate(correct = hits + correct_rejections) %>%
  mutate(incorrect = misses + false_alarms) %>%
  select(gameid, rule_idx, correct, incorrect) %>%
  mutate(explorer_acc = correct / (correct + incorrect))

test_explorer_summary <- test_explorer %>%
  group_by(rule_idx) %>%
  summarize(
    avg_explorer_acc = mean(explorer_acc),
    std_explorer_acc = sd(explorer_acc),
  ) %>%
  select(rule_idx, avg_explorer_acc, std_explorer_acc)

test_student <- scores_complete_games %>%
  filter(role == "student") %>%
  mutate(correct = hits + correct_rejections) %>%
  mutate(incorrect = misses + false_alarms) %>%
  select(gameid, rule_idx, incorrect, correct) %>%
  mutate(student_acc = correct / (correct + incorrect))

test_student_summary <- test_student %>%
  group_by(rule_idx) %>%
  summarize(
    avg_student_acc = mean(student_acc),
    std_student_acc = sd(student_acc),
  ) %>%
  select(rule_idx, avg_student_acc, std_student_acc)

test <- bind_cols(test_explorer, test_student) %>%
    select(rule_idx, student_acc, explorer_acc)
test$rule_idx <- as.factor(test$rule_idx)


test_summary <- bind_cols(test_explorer_summary, test_student_summary) %>%
  select(rule_idx, avg_student_acc, avg_explorer_acc, std_student_acc, std_explorer_acc)
test_summary$rule_idx <- as.factor(test_summary$rule_idx)
```

```{r}
ggplot(test_summary %>%
        rename(student_acc = avg_student_acc) %>%
        rename(explorer_acc = avg_explorer_acc)
       , aes(x=student_acc, y=explorer_acc, color=rule_idx, label=rule_idx)) +
  geom_point(size = 4) +
  theme_bw() +
  guides(color = guide_legend("rule_idx")) +
  geom_text(hjust = 0.01, nudge_x = 0.03, show.legend = FALSE) +
  geom_point(data = test, size = 1) +
  labs(
    title = "Student Acc vs. Explorer Acc",
    x = "Student Acc.",
    y = "Sexplorer Acc."
  ) + geom_segment(aes(x=1, y=1, xend=0.2, yend=0.2), linetype=2, color="black",arrow = arrow(length = unit(0.2, "inches")))
```

## Student versus Teacher Performance -- Pairs (Summurization)
```{r}
ggplot(test_stats_s_unique, aes(x = reorder(role, desc(role)), y = val, fill = stat)) + 
  geom_bar(stat = 'identity', position = 'stack') + facet_grid(~ id)
```

```{r}
test_stats_s_concept = test_stats_s %>%
  mutate(role = as.character(role)) %>%
  mutate(role = replace(role, role == "explorer", "T")) %>%
  mutate(role = replace(role, role == "student", "S")) %>%
  select(role, rule_idx, hits, misses, correct_rejections, false_alarms) %>%
  group_by(rule_idx, role) %>%
  summarize(hits = sum(hits), misses = sum(misses), correct_rejections=sum(correct_rejections), false_alarms=sum(false_alarms)) %>%
  gather(key = "stat", value = "val", 3:6) %>%
  arrange(rule_idx) %>%
  select(rule_idx, role, val, stat)
```
## Student versus Teacher Performance -- Concepts (Summurization)
```{r}
ggplot(test_stats_s_concept, aes(x = reorder(role, desc(role)), y = val, fill = stat)) + 
  geom_bar(stat = 'identity', position = 'stack') + facet_grid(~ rule_idx)  +
  labs(
    title = "Teacher & Student Performance Across Rules (Aggregated)",
    x = "Student & Teachers",
    y = "Counts"
  )
```

```{r}
test_stats_s_rule_type = test_stats_s %>%
  mutate(role = as.character(role)) %>%
  mutate(role = replace(role, role == "explorer", "T")) %>%
  mutate(role = replace(role, role == "student", "S")) %>%
  select(role, rule_type, hits, misses, correct_rejections, false_alarms) %>%
  group_by(rule_type, role) %>%
  summarize(hits = sum(hits), misses = sum(misses), correct_rejections=sum(correct_rejections), false_alarms=sum(false_alarms)) %>%
  gather(key = "stat", value = "val", 3:6) %>%
  select(rule_type, role, val, stat)
```

#-----------------------------------------------
# Hits, Misses, False Alarms, Correct Rejections
# ----------------------------------------------
test_stats_teacher_s <- test_stats_s %>%
  filter(role == "explorer") %>%
  select(rule_idx, hits, misses, correct_rejections, false_alarms) %>%
  filter(rule_idx %in% rule_indices) %>%
  group_by(rule_idx) %>% 
  summarise_all(sum)
summed <- rowSums(test_stats_teacher_s[, c("hits", "misses", "correct_rejections", "false_alarms")])
test_stats_teacher_s_normalized <- test_stats_teacher_s %>%
  mutate(hits = hits / summed) %>%
  mutate(misses = misses / summed) %>%
  mutate(correct_rejections = correct_rejections / summed) %>%
  mutate(false_alarms = false_alarms / summed)

test_stats_student_s <- test_stats_s %>%
  filter(role == "student") %>%
  select(rule_idx, hits, misses, correct_rejections, false_alarms) %>%
  filter(rule_idx %in% rule_indices) %>%
  group_by(rule_idx) %>% 
  summarise_all(sum)
summed <- rowSums(test_stats_student_s[, c("hits", "misses", "correct_rejections", "false_alarms")])
test_stats_student_s_normalized <- test_stats_student_s %>%
  mutate(hits = hits / summed) %>%
  mutate(misses = misses / summed) %>%
  mutate(correct_rejections = correct_rejections / summed) %>%
  mutate(false_alarms = false_alarms / summed)



## Student versus Teacher Performance -- Concept Types (Summurization)
```{r fig.width=7, fig.height=4, echo=FALSE}
ggplot(test_stats_s_rule_type, aes(x = reorder(role, desc(role)), y = val, fill = stat)) + 
  geom_bar(stat = 'identity', position = 'stack') + facet_grid(~ rule_type) +
    labs(
    title = "Teacher & Student Performance Across Concept Types (Aggregated)",
    x = "Student & Teachers",
    y = "Counts"
  )
```

## Time Comparisons (Summarization)
```{r}
train_times <- train_trials_s %>%
    select(rule_idx, total_time) %>%
    filter(rule_idx %in% rule_indices) %>%
    group_by(rule_idx) %>% 
    summarise_all(mean) %>%
    mutate(role = "T") %>%
    rename(train_time = total_time)

test_times <- bind_rows(
  test_stats_s %>%
    filter(role == "explorer") %>%
    select(rule_idx, total_time) %>%
    filter(rule_idx %in% rule_indices) %>%
    group_by(rule_idx) %>% 
    summarise_all(mean) %>%
    mutate(role = "T"),
  
  test_stats_s %>%
    filter(role == "student") %>%
    select(rule_idx, total_time) %>%
    filter(rule_idx %in% rule_indices) %>%
    group_by(rule_idx) %>% 
    summarise_all(mean) %>%
    mutate(role = "S")
) %>%
  rename(test_time = total_time)

chat_times <- chat_messages %>%
    filter(rule_idx %in% rule_indices) %>%
    select(rule_idx, gameid, reactionTime) %>%
    group_by(rule_idx, gameid) %>%
    summarize(reactionTime = sum(reactionTime)) %>%
    group_by(rule_idx) %>%
    summarize(reactionTime = mean(reactionTime) / 1000) %>%
    rename(chat_time = reactionTime)

train_chat_times <- bind_cols(train_times, chat_times) %>%
  select(rule_idx, train_time, chat_time, role) %>%
  gather(key = "stat", value = "val", 2:3) %>%
  select(rule_idx, role, val, stat) %>%
```


```{r fig.width=7, fig.height=4, echo=FALSE}
ggplot(train_chat_times, aes(x = reorder(stat, desc(stat)), y = val)) + 
  geom_bar(stat = 'identity') + facet_grid(~ rule_idx) +
    labs(
    title = "Train & Chat Time Comparisons",
    x = "Student & Teachers",
    y = "Time (Seconds)"
  )
```

For Simple Rules 0, 2, 4 (Single Feature & Conjunction) we see that there isn't a wide variance in terms of train time and chat time; but this rift becomes noticeable when we have Harder Rules 6, 8 (Disjunction).

# Corpus Analysis

## Breakdown of Dialogue Statements (Categories)
```{r}
chat_messages_count = chat_messages %>%
  mutate(role = as.character(role)) %>%
  mutate(role = replace(role, role == "explorer", "T")) %>%
  mutate(role = replace(role, role == "student", "S")) %>%
  select(role, messageType) %>%
  group_by(role) %>%
  summarize(
    informative = sum(messageType == 'Informative'),
    follow_up = sum(messageType == 'Follow-Up'),
    social = sum(messageType == 'Social'),
    misc = sum(messageType == 'Misc'),
    confirmation = sum(messageType == 'Confirmation'),
  ) %>%
  select(role, informative, follow_up, social, misc, confirmation) %>%
  gather(key = "stat", value = "val", 2:6) %>%
  select(role, val, stat) %>%
  group_by(role) %>%
  mutate(val_percentage_of_role = val/sum(val))
```
```{r}
ggplot(chat_messages_count, aes(x = reorder(role, desc(role)), y = val_percentage_of_role, fill = stat)) + 
  geom_bar(stat = 'identity', position = 'stack') + ggtitle("% of dialogue messages produced by each player") + xlab('Role') + ylab('% of player\'s messages')
```

```{r}
ggplot(chat_messages_count, aes(x = reorder(role, desc(role)), y = val, fill = stat)) + 
  geom_bar(stat = 'identity', position = 'stack') + ggtitle("# of messages produced by each player") +
  xlab('Role') + ylab('# of player\'s messages')
```



```{r}
ggplot(
    chat_messages_count %>%
      select(val, stat) %>%
      group_by(stat) %>%
      summarize(val = sum(val)),
    aes(x = stat, y = val)
  ) + geom_bar(stat = 'identity') + ggtitle("# of Messages, Across Type") +
    xlab('Message Type') + ylab('# of Messages Produced')
```



```{r}
chat_messages_word_count = chat_messages %>%
  mutate(role = as.character(role)) %>%
  mutate(role = replace(role, role == "explorer", "T")) %>%
  mutate(role = replace(role, role == "student", "S")) %>%
  select(role, messageType, text) %>%
  group_by(role, messageType) %>%
  summarise(
    messageWordLength = sum(sapply(strsplit(text, " "), length))
  )
```

```{r}
  ggplot(chat_messages_word_count, aes(x = reorder(role, desc(role)), y = messageWordLength, fill = messageType)) + 
  geom_bar(stat = 'identity', position = 'stack') + 
  ggtitle("# of words produced by each player, according to utterance category") + 
  xlab('Role') + ylab('# of player\'s words')
```


## Informative Message Types
```{r}
informative_messages_count = chat_messages %>%
  filter(messageType == 'Informative') %>%
  select(utteranceType) %>%
  summarize(
    None = sum(utteranceType == 'None'),
    Quantifier = sum(utteranceType == 'Quantifier'),
    Conditional = sum(utteranceType == 'Conditional'),
    Exemplar = sum(utteranceType == 'Exemplar'),
    Generic = sum(utteranceType == 'Generic'),
    Imperative = sum(utteranceType == 'Imperative'),
    Affirmative = sum(utteranceType == 'Affirmative'),
    Knowledge = sum(utteranceType == 'Knowledge'),
    Logic = sum(utteranceType == 'Logic'),
    Question = sum(utteranceType == 'Question'),
    Possibility_Conditional = sum(utteranceType == 'Possibility,Conditional'),
  ) %>%
  select(None, Quantifier, Conditional, Exemplar, Generic, Imperative, Affirmative, Knowledge, Logic, Question, Possibility_Conditional) %>%
  gather(key = "stat", value = "val")
```

```{r fig.width=7, fig.height=4, echo=FALSE}
ggplot(
    informative_messages_count,
    aes(x = stat, y = val)
  ) + geom_bar(stat = 'identity') + ggtitle("# of Informative Messages, Across Utterance Type") +
    xlab('Utterance Type') + ylab('# of Messages Produced')
```

```{r}
informative_messages_word_count = chat_messages %>%
  filter(messageType == 'Informative') %>%
  select(utteranceType, text) %>%
  group_by(utteranceType) %>%
  summarise(
    messageWordLength = sum(sapply(strsplit(text, " "), length))
  )
```

```{r fig.width=7, fig.height=4, echo=FALSE}
ggplot(
    informative_messages_word_count,
    aes(x = utteranceType, y = messageWordLength)
  ) + geom_bar(stat = 'identity') + ggtitle("# of Informative Words, Across Utterance Type") +
    xlab('Utterance Type') + ylab('# of Words Produced')
```


## Follow-Up Message Types
```{r}
follow_up_messages_count = chat_messages %>%
  filter(messageType == 'Follow-Up') %>%
  mutate(role = as.character(role)) %>%
  mutate(role = replace(role, role == "explorer", "T")) %>%
  mutate(role = replace(role, role == "student", "S")) %>%
  group_by(role) %>%
  select(role, utteranceType) %>%
  summarize(
    None = sum(utteranceType == 'None'),
    Quantifier = sum(utteranceType == 'Quantifier'),
    Conditional = sum(utteranceType == 'Conditional'),
    Exemplar = sum(utteranceType == 'Exemplar'),
    Generic = sum(utteranceType == 'Generic'),
    Imperative = sum(utteranceType == 'Imperative'),
    Affirmative = sum(utteranceType == 'Affirmative'),
    Knowledge = sum(utteranceType == 'Knowledge'),
    Logic = sum(utteranceType == 'Logic'),
    Question = sum(utteranceType == 'Question'),
    Possibility_Conditional = sum(utteranceType == 'Possibility,Conditional'),
  ) %>%
  select(role, None, Quantifier, Conditional, Exemplar, Generic, Imperative, Affirmative, Knowledge, Logic, Question, Possibility_Conditional) %>%
  gather(key = "stat", value = "val", -role)
```


```{r fig.width=7, fig.height=4, echo=FALSE}
ggplot(
    follow_up_messages_count %>%
      select(val, stat) %>%
      group_by(stat) %>%
      summarize(val = sum(val)),
    aes(x = stat, y = val)
  ) + geom_bar(stat = 'identity') + ggtitle("# of Follow-Up Messages, Across Utterance Type") +
    xlab('Message Type') + ylab('# of Messages Produced')
```

```{r fig.width=7, fig.height=4, echo=FALSE}
ggplot(
    follow_up_messages_count,
    aes(x = reorder(role, desc(role)), y = val, fill=stat)
  ) + geom_bar(stat = 'identity', position = 'stack') + ggtitle("# of Follow-Up Messages, Across Utterance Type") +
    xlab('Utterance Type') + ylab('# of Messages Produced')
```

## Quantifier Specific Language
```{r}
quantifier_messages_count = chat_messages %>%
  filter(
    utteranceType == 'Quantifier' | 
      (utteranceType == 'Question' & 
         (grepl('any', text)  | 
            grepl('all', text) |
            grepl('mostly', text) | 
            grepl('some', text)
          )
       )
  ) %>%
  mutate(utteranceTypeSignifiers=replace(utteranceTypeSignifiers, utteranceType=='Question', 'Question')) %>%
  select(role, utteranceTypeSignifiers) %>%
  group_by(role, utteranceTypeSignifiers) %>%
  summarize(
    Count = length(utteranceTypeSignifiers)
  )
```

```{r fig.width=7, fig.height=4, echo=FALSE}
ggplot(
    quantifier_messages_count,
    aes(x = utteranceTypeSignifiers, y = Count)
  ) + geom_bar(stat = 'identity') + ggtitle("# of Messages with Quantifier Signifier (Teacher & Student Combined)") +
    xlab('Signifier') + ylab('# of Times Utilized')
```


```{r fig.width=7, fig.height=4, echo=FALSE}
ggplot(
    quantifier_messages_count,
    aes(x = role, y = Count, fill=utteranceTypeSignifiers)
  ) + geom_bar(stat = 'identity', position ='stack') + ggtitle("# of Messages with Quantifier Signifier (Teacher & Student Split)") +
    xlab('Role') + ylab('# of Times Utilized')
```

## Analysis of Langauge Across Pairs
```{r}
game_informative_msgs = chat_messages %>%
  filter(messageType == 'Informative') %>%
  group_by(gameid) %>%
  select(gameid, utteranceType) %>%
  summarize(
    None = sum(utteranceType == 'None'),
    Quantifier = sum(utteranceType == 'Quantifier'),
    Conditional = sum(utteranceType == 'Conditional'),
    Exemplar = sum(utteranceType == 'Exemplar'),
    Generic = sum(utteranceType == 'Generic'),
    Imperative = sum(utteranceType == 'Imperative'),
    Affirmative = sum(utteranceType == 'Affirmative'),
    Knowledge = sum(utteranceType == 'Knowledge'),
    Logic = sum(utteranceType == 'Logic'),
    Question = sum(utteranceType == 'Question'),
    Possibility_Conditional = sum(utteranceType == 'Possibility,Conditional'),
  ) %>%
  select(gameid, None, Quantifier, Conditional, Exemplar, Generic, Imperative, Affirmative, Knowledge, Logic, Question, Possibility_Conditional) %>%
  gather(key = "stat", value = "val", -gameid)
```

```{r fig.width=7, fig.height=4, echo=FALSE}
ggplot(
    game_informative_msgs,
    aes(x='', y=val, fill=stat)
  ) + geom_bar(stat = 'identity', position ='stack') + ggtitle("Distribution of Utterance Types (Informative) Across Teachers") +
    xlab('Teacher') + ylab('# of Messages of Utterance Type') + facet_grid(~gameid)
```

```{r}
pair_messages = chat_messages %>%
  filter(messageType == 'Follow-Up' | messageType == 'Informative') %>%
  mutate(role = as.character(role)) %>%
  mutate(role = replace(role, role == "explorer", "T")) %>%
  mutate(role = replace(role, role == "student", "S")) %>%
  group_by(gameid, role) %>%
  select(role, gameid, utteranceType) %>%
  summarize(
    None = sum(utteranceType == 'None'),
    Quantifier = sum(utteranceType == 'Quantifier'),
    Conditional = sum(utteranceType == 'Conditional'),
    Exemplar = sum(utteranceType == 'Exemplar'),
    Generic = sum(utteranceType == 'Generic'),
    Imperative = sum(utteranceType == 'Imperative'),
    Affirmative = sum(utteranceType == 'Affirmative'),
    Knowledge = sum(utteranceType == 'Knowledge'),
    Logic = sum(utteranceType == 'Logic'),
    Question = sum(utteranceType == 'Question'),
    Possibility_Conditional = sum(utteranceType == 'Possibility,Conditional'),
  ) %>%
  select(role, gameid, None, Quantifier, Conditional, Exemplar, Generic, Imperative, Affirmative, Knowledge, Logic, Question, Possibility_Conditional) %>%
  gather(key = "stat", value = "val", 3:13)
```

```{r fig.width=7, fig.height=4, echo=FALSE}
ggplot(
    pair_messages,
    aes(x=role, y=val, fill=stat)
  ) + geom_bar(stat = 'identity', position ='stack') + ggtitle("Distribution of Utterance Types (Informative + Follow-Up) Across Player Pairs ") +
    xlab('Role') + ylab('# of Messages of Utterance Type') + facet_grid(~gameid)
```

```{r}
ggplot(y,
    aes(x = rule_idx, y = n)
) + geom_bar(stat = 'identity') + ggtitle("# of Games By Rule Index") +
    xlab('Rule Index') + ylab('# Games Played')
```

## Analysis of Language Across Concepts

## Analysis of Language Across Concept Types 


