---
title: "Cultural Ratchet Data Exploration"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(reshape2)
library(purrr)
library(jsonlite)
library(tidyr) 
library(dplyr)
library(tidyboot)
library(ggplot2)
library(scales)
library(stringr)
library(digest)
```

```{r}
#-----------------------
# Reading Mp-Game-6 Data
#-----------------------

# All Games

# Load Test Trial Responses
temp <- list.files(
  "../../../data/mp-game-6/all_games/logTest",
  pattern="*.csv",
  full.names=TRUE
)
test_all_games <- do.call(rbind, lapply(temp, read.csv, sep=""))

# Load Test Summary Stats
temp <- list.files(
    "../../../data/mp-game-6/all_games/logScores",
  pattern="*.csv",
  full.names=TRUE
)
scores_all_games <- do.call(rbind, lapply(temp, read.csv, sep=""))

# Load Chat Messages
temp <- list.files(
    "../../../data/mp-game-6/all_games/chatMessage",
  pattern="*.csv",
  full.names=TRUE
)
msgs_complete_games <-do.call(rbind, lapply(temp, readr::read_tsv)) %>%
  group_by(gameid)

# Complete Games

# Load Test Trial Responses
temp <- list.files(
  "../../../data/mp-game-6/complete_games/logTest",
  pattern="*.csv",
  full.names=TRUE
)
test_complete_games <- do.call(rbind, lapply(temp, read.csv, sep=""))

# Load Test Summary Stats
temp <- list.files(
    "../../../data/mp-game-6/complete_games/logScores",
  pattern="*.csv",
  full.names=TRUE
)
scores_complete_games <- do.call(rbind, lapply(temp, read.csv, sep=""))

# Load Chat Messages
temp <- list.files(
    "../../../data/mp-game-6/complete_games/chatMessage",
  pattern="*.csv",
  full.names=TRUE
)
msgs_complete_games <-do.call(rbind, lapply(temp, readr::read_tsv)) %>%
  group_by(gameid)
```

# Distribution of Complete Data
```{r}
num_complete_games_played <- scores_complete_games %>%
  count(rule_idx) %>%
  mutate(n = n/2)
num_complete_games_played$n <- as.factor(num_complete_games_played$n)
ggplot(data=num_complete_games_played, aes(x=rule_idx, y=n)) +
  geom_bar(stat="identity") +
  labs(
    title = "Data Distribution of Complete Games",
    x = "Rule Index",
    y = "Number of Samples"
  )  +
  ylim(0, 10) +
  scale_y_discrete()
```

# Distribution of All Data
```{r}
num_all_games_played <- scores_all_games %>%
  count(rule_idx) %>%
  mutate(n = ceiling(n/2))
num_all_games_played$n <- as.factor(num_all_games_played$n)
ggplot(data=num_all_games_played, aes(x=rule_idx, y=n), ) +
  geom_bar(stat="identity") +
  labs(
    title = "Data Distribution of All Games",
    x = "Rule Index",
    y = "Number of Samples"
  ) +
  ylim(0, 10) +
  scale_y_discrete()
```

# Load Concept Summary
```{r}
concept_summary_fp <- c("../../../experiments/mp-game-6/stimuli/fifty_rules/concept_summary.json")
concept_summary <- do.call(rbind.data.frame, jsonlite::fromJSON(concept_summary_fp)) %>%
  mutate(chance_rate= p_test_belongs_to_concept*p_test_belongs_to_concept + (1 -p_test_belongs_to_concept) * (1-p_test_belongs_to_concept))
```


# Student Performance Vs. Teacher Performance (Complete Game)
```{r}
# Accuracies
complete_test_explorer <- scores_complete_games %>%
  filter(role == "explorer") %>%
  mutate(correct = hits + correct_rejections) %>%
  mutate(incorrect = misses + false_alarms) %>%
  select(gameid, rule_idx, correct, incorrect) %>%
  mutate(explorer_acc = correct / (correct + incorrect)) %>%
  mutate(rule_type = floor(rule_idx/10))

complete_test_student <- scores_complete_games %>%
  filter(role == "student") %>%
  mutate(correct = hits + correct_rejections) %>%
  mutate(incorrect = misses + false_alarms) %>%
  select(gameid, rule_idx, incorrect, correct) %>%
  mutate(student_acc = correct / (correct + incorrect)) %>%
  mutate(rule_type = floor(rule_idx/10))

complete_test_explorer_summary <- complete_test_explorer %>%
  group_by(rule_idx) %>%
  tidyboot_mean(column = explorer_acc) %>%
  rename(avg_explorer_acc = mean) %>%
  rename(explorer_acc_ci_lower = ci_lower) %>%
  rename(explorer_acc_ci_upper = ci_upper )%>%
  mutate(rule_type = floor(rule_idx/10))

complete_test_student_summary <- complete_test_student %>%
  group_by(rule_idx) %>%
  tidyboot_mean(column = student_acc) %>%
  rename(avg_student_acc = mean) %>%
  rename(student_acc_ci_lower = ci_lower) %>%
  rename(student_acc_ci_upper = ci_upper) %>%
  mutate(rule_type = floor(rule_idx/10))

complete_test <- bind_cols(complete_test_explorer, complete_test_student) %>%
    select(rule_idx, student_acc, explorer_acc, rule_type) 
complete_test$rule_idx <- as.factor(complete_test$rule_idx)

complete_test_summary <- bind_cols(complete_test_explorer_summary, complete_test_student_summary) %>% bind_cols(concept_summary) %>%
  select(rule_idx, avg_student_acc, avg_explorer_acc, student_acc_ci_lower, student_acc_ci_upper, explorer_acc_ci_lower,  explorer_acc_ci_upper, p_test_belongs_to_concept, phrase, kind, name, type, rule_type, chance_rate) 
complete_test_summary$rule_idx <- as.factor(complete_test_summary$rule_idx)
complete_test_summary$type <- as.factor(complete_test_summary$type)
```

```{r}
ggplot(complete_test_summary %>%
        rename(student_acc = avg_student_acc) %>%
        rename(explorer_acc = avg_explorer_acc),
       aes(x=student_acc, y=explorer_acc, color=rule_type, label=rule_idx))+
  geom_point(size = 2) +
  theme_bw() +
  guides(color = guide_legend("rule_type")) +
  geom_text(hjust = 0.01, nudge_x = 0.01, show.legend = FALSE) +
  geom_point(data = complete_test, size = 1) +
  labs(
    title = "Student Acc vs. Explorer Acc",
    x = "Student Acc.",
    y = "Explorer Acc."
  ) + geom_segment(aes(x=1, y=1, xend=0.2, yend=0.2), linetype=2, color="black",arrow = arrow(length = unit(0.2, "inches")))
```
# Student Performance Vs. Teacher Performance (All Games)
```{r}
# Accuracies
all_test_explorer <- scores_all_games %>%
  filter(role == "explorer") %>%
  mutate(correct = hits + correct_rejections) %>%
  mutate(incorrect = misses + false_alarms) %>%
  select(gameid, rule_idx, correct, incorrect) %>%
  mutate(explorer_acc = correct / (correct + incorrect)) %>%
  mutate(rule_type = floor(rule_idx/10))


all_test_student <- scores_all_games %>%
  filter(role == "student") %>%
  mutate(correct = hits + correct_rejections) %>%
  mutate(incorrect = misses + false_alarms) %>%
  select(gameid, rule_idx, incorrect, correct) %>%
  mutate(student_acc = correct / (correct + incorrect)) %>%
  mutate(rule_type = floor(rule_idx/10))

all_test <- bind_cols(all_test_explorer, all_test_student) %>%
    select(rule_idx, student_acc, explorer_acc, rule_type)
all_test$rule_idx <- as.factor(all_test$rule_idx)
all_test$rule_type <- as.factor(all_test$rule_type)

all_test_explorer_summary <- all_test_explorer %>%
  group_by(rule_idx) %>%
  tidyboot_mean(column = explorer_acc) %>%
  rename(avg_explorer_acc = mean) %>%
  rename(explorer_acc_ci_lower = ci_lower) %>%
  rename(explorer_acc_ci_upper = ci_upper )%>%
  mutate(rule_type = floor(rule_idx/10))

all_test_student_summary <- all_test_student %>%
  group_by(rule_idx) %>%
  tidyboot_mean(column = student_acc) %>%
  rename(avg_student_acc = mean) %>%
  rename(student_acc_ci_lower = ci_lower) %>%
  rename(student_acc_ci_upper = ci_upper) %>%
  mutate(rule_type = floor(rule_idx/10))


all_test_summary <- bind_cols(all_test_explorer_summary, all_test_student_summary) %>%
  bind_cols(concept_summary) %>%
  select(rule_idx, avg_student_acc, avg_explorer_acc, student_acc_ci_lower, student_acc_ci_upper, explorer_acc_ci_lower, explorer_acc_ci_upper, rule_type, p_test_belongs_to_concept, phrase, kind, name, type) %>%
  mutate(chance_rate= p_test_belongs_to_concept*p_test_belongs_to_concept + (1 -p_test_belongs_to_concept) * (1-p_test_belongs_to_concept))
all_test_summary$rule_idx <- as.factor(all_test_summary$rule_idx)
all_test_summary$rule_type <- as.factor(all_test_summary$rule_type)
```

```{r}
ggplot(all_test_summary %>%
        rename(student_acc = avg_student_acc) %>%
        rename(explorer_acc = avg_explorer_acc)
       , aes(x=student_acc, y=explorer_acc, color=rule_type, label=rule_idx)) +
  geom_point(size = 2) +
  theme_bw() +
  guides(color = guide_legend("Rule Type")) +
  geom_text(hjust = 0.01, nudge_x = 0.01, show.legend = FALSE) +
  geom_point(data = all_test, size = 0.5) +
  labs(
    title = "Student Acc vs. Explorer Acc",
    x = "Student Acc.",
    y = "Explorer Acc."
  ) + geom_segment(aes(x=1, y=1, xend=0.2, yend=0.2), linetype=2, color="black",arrow = arrow(length = unit(0.2, "inches"))) +
  geom_errorbar(aes(ymax = explorer_acc_ci_lower, ymin = explorer_acc_ci_upper)) +
  geom_errorbarh(aes(xmax = student_acc_ci_lower, xmin = student_acc_ci_upper))
```
```{r}
ggplot(all_test_summary %>%
        rename(student_acc = avg_student_acc) %>%
        rename(explorer_acc = avg_explorer_acc)
       , aes(x=student_acc, y=explorer_acc, color=rule_type, label=rule_idx)) +
  geom_point(size = 2) +
  theme_bw() +
  guides(color = guide_legend("Rule Type")) +
  geom_text(hjust = 0.01, nudge_x = 0.01, show.legend = FALSE) +
  geom_point(data = all_test, size = 0.5) +
  labs(
    title = "Student Acc vs. Explorer Acc",
    x = "Student Acc.",
    y = "Explorer Acc."
  ) + geom_segment(aes(x=1, y=1, xend=0.2, yend=0.2), linetype=2, color="black",arrow = arrow(length = unit(0.2, "inches"))) +
  geom_errorbar(aes(ymax = explorer_acc_ci_lower, ymin = explorer_acc_ci_upper)) +
  geom_errorbarh(aes(xmax = student_acc_ci_lower, xmin = student_acc_ci_upper))
```

```{r}
# Plot of base rates and performance of students and teachers
ggplot(all_test_summary %>%
        rename(student_acc = avg_student_acc) %>%
        rename(explorer_acc = avg_explorer_acc)
       , aes(x=rule_idx, y=explorer_acc)) + 
  geom_point(col="red", size=) +   # Draw points
  geom_point(aes(y=chance_rate), col="black", size=3) +
  geom_point(aes(y=student_acc), col="blue", size=2) + 
  coord_flip() + 
  theme_bw() +
  labs( 
    title="Chance Vs. Mean Participant Performance", 
    y="Accuracy (%)", 
    x="Rule Index")
```

### Concept Type Bar Plot
```{r}
all_test_concept_type_summary <- scores_all_games %>%
  mutate(correct = hits + correct_rejections) %>%
  mutate(incorrect = misses + false_alarms) %>%
  select(gameid, rule_idx, correct, incorrect, role) %>%
  mutate(acc = correct / (correct + incorrect)) %>%
  mutate(rule_type = floor(rule_idx/10))  %>%
  group_by(rule_type, role) %>%
  tidyboot_mean(column = acc)

all_test_concept_type_summary$rule_type <- as.factor(all_test_concept_type_summary$rule_type)
all_test_concept_type_summary$role <- as.factor(all_test_concept_type_summary$role)
```

```{r fig.width=7, fig.height=4, echo=FALSE}
ggplot(all_test_concept_type_summary, aes(rule_type, fill=role, y=empirical_stat)) + 
  geom_col(position='dodge') +  
  labs(
    title = "Concept Type Performance",
    x = "Concept Type",
    y = "Accuracy %"
  ) + 
  geom_errorbar(aes(x=rule_type, ymin=ci_lower, ymax=ci_upper), color="black", width=0.5, position=position_dodge(.9)) 
```

### Concept Type Bar Plot -- Normalized Accuracies 
```{r}
all_test_concept_type_summary_normalized <- scores_all_games %>%
  mutate(correct = hits + correct_rejections) %>%
  mutate(incorrect = misses + false_alarms) %>%
  select(gameid, rule_idx, correct, incorrect, role) %>%
  mutate(acc = correct / (correct + incorrect)) %>%
  mutate(rule_type = floor(rule_idx/10)) %>%
  mutate(chance_rate = concept_summary$chance_rate[rule_idx + 1]) %>%
  mutate(norm_acc = (acc - chance_rate)/(1 - chance_rate)) %>%
  group_by(rule_type, role) %>%
  tidyboot_mean(column = norm_acc)

all_test_concept_type_summary_normalized$rule_type <- as.factor(all_test_concept_type_summary_normalized$rule_type)
all_test_concept_type_summary_normalized$role <- as.factor(all_test_concept_type_summary_normalized$role)
```

```{r fig.width=7, fig.height=4, echo=FALSE}
ggplot(all_test_concept_type_summary_normalized, aes(rule_type, fill=role, y=empirical_stat)) + 
  geom_col(position='dodge') +  
  labs(
    title = "Concept Type Performance",
    x = "Concept Type",
    y = "Accuracy %"
  ) + 
  geom_errorbar(aes(x=rule_type, ymin=ci_lower, ymax=ci_upper), color="black", width=0.5, position=position_dodge(.9)) 
```

```{r}
#-----------------------------------------------
# Hits, Misses, False Alarms, Correct Rejections
# ----------------------------------------------
test_stats_teacher_s <- test_stats_s %>%
  filter(role == "explorer") %>%
  select(rule_idx, hits, misses, correct_rejections, false_alarms) %>%
  filter(rule_idx %in% rule_indices) %>%
  group_by(rule_idx) %>% 
  summarise_all(sum)
summed <- rowSums(test_stats_teacher_s[, c("hits", "misses", "correct_rejections", "false_alarms")])
test_stats_teacher_s_normalized <- test_stats_teacher_s %>%
  mutate(hits = hits / summed) %>%
  mutate(misses = misses / summed) %>%
  mutate(correct_rejections = correct_rejections / summed) %>%
  mutate(false_alarms = false_alarms / summed)

test_stats_student_s <- test_stats_s %>%
  filter(role == "student") %>%
  select(rule_idx, hits, misses, correct_rejections, false_alarms) %>%
  filter(rule_idx %in% rule_indices) %>%
  group_by(rule_idx) %>% 
  summarise_all(sum)
summed <- rowSums(test_stats_student_s[, c("hits", "misses", "correct_rejections", "false_alarms")])
test_stats_student_s_normalized <- test_stats_student_s %>%
  mutate(hits = hits / summed) %>%
  mutate(misses = misses / summed) %>%
  mutate(correct_rejections = correct_rejections / summed) %>%
  mutate(false_alarms = false_alarms / summed)
```


## Student versus Teacher Performance -- Concept Types (Summurization)
```{r fig.width=7, fig.height=4, echo=FALSE}
ggplot(test_stats_s_rule_type, aes(x = reorder(role, desc(role)), y = val, fill = stat)) + 
  geom_bar(stat = 'identity', position = 'stack') + facet_grid(~ rule_type) +
    labs(
    title = "Teacher & Student Performance Across Concept Types (Aggregated)",
    x = "Student & Teachers",
    y = "Counts"
  )
```

## Time Comparisons (Summarization)
```{r}
train_times <- train_trials_s %>%
    select(rule_idx, total_time) %>%
    filter(rule_idx %in% rule_indices) %>%
    group_by(rule_idx) %>% 
    summarise_all(mean) %>%
    mutate(role = "T") %>%
    rename(train_time = total_time)

test_times <- bind_rows(
  test_stats_s %>%
    filter(role == "explorer") %>%
    select(rule_idx, total_time) %>%
    filter(rule_idx %in% rule_indices) %>%
    group_by(rule_idx) %>% 
    summarise_all(mean) %>%
    mutate(role = "T"),
  
  test_stats_s %>%
    filter(role == "student") %>%
    select(rule_idx, total_time) %>%
    filter(rule_idx %in% rule_indices) %>%
    group_by(rule_idx) %>% 
    summarise_all(mean) %>%
    mutate(role = "S")
) %>%
  rename(test_time = total_time)

chat_times <- chat_messages %>%
    filter(rule_idx %in% rule_indices) %>%
    select(rule_idx, gameid, reactionTime) %>%
    group_by(rule_idx, gameid) %>%
    summarize(reactionTime = sum(reactionTime)) %>%
    group_by(rule_idx) %>%
    summarize(reactionTime = mean(reactionTime) / 1000) %>%
    rename(chat_time = reactionTime)

train_chat_times <- bind_cols(train_times, chat_times) %>%
  select(rule_idx, train_time, chat_time, role) %>%
  gather(key = "stat", value = "val", 2:3) %>%
  select(rule_idx, role, val, stat) %>%
```


```{r fig.width=7, fig.height=4, echo=FALSE}
ggplot(train_chat_times, aes(x = reorder(stat, desc(stat)), y = val)) + 
  geom_bar(stat = 'identity') + facet_grid(~ rule_idx) +
    labs(
    title = "Train & Chat Time Comparisons",
    x = "Student & Teachers",
    y = "Time (Seconds)"
  )
```

For Simple Rules 0, 2, 4 (Single Feature & Conjunction) we see that there isn't a wide variance in terms of train time and chat time; but this rift becomes noticeable when we have Harder Rules 6, 8 (Disjunction).

# Corpus Analysis

## Breakdown of Dialogue Statements (Categories)
```{r}
chat_messages_count = chat_messages %>%
  mutate(role = as.character(role)) %>%
  mutate(role = replace(role, role == "explorer", "T")) %>%
  mutate(role = replace(role, role == "student", "S")) %>%
  select(role, messageType) %>%
  group_by(role) %>%
  summarize(
    informative = sum(messageType == 'Informative'),
    follow_up = sum(messageType == 'Follow-Up'),
    social = sum(messageType == 'Social'),
    misc = sum(messageType == 'Misc'),
    confirmation = sum(messageType == 'Confirmation'),
  ) %>%
  select(role, informative, follow_up, social, misc, confirmation) %>%
  gather(key = "stat", value = "val", 2:6) %>%
  select(role, val, stat) %>%
  group_by(role) %>%
  mutate(val_percentage_of_role = val/sum(val))
```
```{r}
ggplot(chat_messages_count, aes(x = reorder(role, desc(role)), y = val_percentage_of_role, fill = stat)) + 
  geom_bar(stat = 'identity', position = 'stack') + ggtitle("% of dialogue messages produced by each player") + xlab('Role') + ylab('% of player\'s messages')
```

```{r}
ggplot(chat_messages_count, aes(x = reorder(role, desc(role)), y = val, fill = stat)) + 
  geom_bar(stat = 'identity', position = 'stack') + ggtitle("# of messages produced by each player") +
  xlab('Role') + ylab('# of player\'s messages')
```



```{r}
ggplot(
    chat_messages_count %>%
      select(val, stat) %>%
      group_by(stat) %>%
      summarize(val = sum(val)),
    aes(x = stat, y = val)
  ) + geom_bar(stat = 'identity') + ggtitle("# of Messages, Across Type") +
    xlab('Message Type') + ylab('# of Messages Produced')
```



```{r}
chat_messages_word_count = chat_messages %>%
  mutate(role = as.character(role)) %>%
  mutate(role = replace(role, role == "explorer", "T")) %>%
  mutate(role = replace(role, role == "student", "S")) %>%
  select(role, messageType, text) %>%
  group_by(role, messageType) %>%
  summarise(
    messageWordLength = sum(sapply(strsplit(text, " "), length))
  )
```

```{r}
  ggplot(chat_messages_word_count, aes(x = reorder(role, desc(role)), y = messageWordLength, fill = messageType)) + 
  geom_bar(stat = 'identity', position = 'stack') + 
  ggtitle("# of words produced by each player, according to utterance category") + 
  xlab('Role') + ylab('# of player\'s words')
```


## Informative Message Types
```{r}
informative_messages_count = chat_messages %>%
  filter(messageType == 'Informative') %>%
  select(utteranceType) %>%
  summarize(
    None = sum(utteranceType == 'None'),
    Quantifier = sum(utteranceType == 'Quantifier'),
    Conditional = sum(utteranceType == 'Conditional'),
    Exemplar = sum(utteranceType == 'Exemplar'),
    Generic = sum(utteranceType == 'Generic'),
    Imperative = sum(utteranceType == 'Imperative'),
    Affirmative = sum(utteranceType == 'Affirmative'),
    Knowledge = sum(utteranceType == 'Knowledge'),
    Logic = sum(utteranceType == 'Logic'),
    Question = sum(utteranceType == 'Question'),
    Possibility_Conditional = sum(utteranceType == 'Possibility,Conditional'),
  ) %>%
  select(None, Quantifier, Conditional, Exemplar, Generic, Imperative, Affirmative, Knowledge, Logic, Question, Possibility_Conditional) %>%
  gather(key = "stat", value = "val")
```

```{r fig.width=7, fig.height=4, echo=FALSE}
ggplot(
    informative_messages_count,
    aes(x = stat, y = val)
  ) + geom_bar(stat = 'identity') + ggtitle("# of Informative Messages, Across Utterance Type") +
    xlab('Utterance Type') + ylab('# of Messages Produced')
```

```{r}
informative_messages_word_count = chat_messages %>%
  filter(messageType == 'Informative') %>%
  select(utteranceType, text) %>%
  group_by(utteranceType) %>%
  summarise(
    messageWordLength = sum(sapply(strsplit(text, " "), length))
  )
```

```{r fig.width=7, fig.height=4, echo=FALSE}
ggplot(
    informative_messages_word_count,
    aes(x = utteranceType, y = messageWordLength)
  ) + geom_bar(stat = 'identity') + ggtitle("# of Informative Words, Across Utterance Type") +
    xlab('Utterance Type') + ylab('# of Words Produced')
```


## Follow-Up Message Types
```{r}
follow_up_messages_count = chat_messages %>%
  filter(messageType == 'Follow-Up') %>%
  mutate(role = as.character(role)) %>%
  mutate(role = replace(role, role == "explorer", "T")) %>%
  mutate(role = replace(role, role == "student", "S")) %>%
  group_by(role) %>%
  select(role, utteranceType) %>%
  summarize(
    None = sum(utteranceType == 'None'),
    Quantifier = sum(utteranceType == 'Quantifier'),
    Conditional = sum(utteranceType == 'Conditional'),
    Exemplar = sum(utteranceType == 'Exemplar'),
    Generic = sum(utteranceType == 'Generic'),
    Imperative = sum(utteranceType == 'Imperative'),
    Affirmative = sum(utteranceType == 'Affirmative'),
    Knowledge = sum(utteranceType == 'Knowledge'),
    Logic = sum(utteranceType == 'Logic'),
    Question = sum(utteranceType == 'Question'),
    Possibility_Conditional = sum(utteranceType == 'Possibility,Conditional'),
  ) %>%
  select(role, None, Quantifier, Conditional, Exemplar, Generic, Imperative, Affirmative, Knowledge, Logic, Question, Possibility_Conditional) %>%
  gather(key = "stat", value = "val", -role)
```


```{r fig.width=7, fig.height=4, echo=FALSE}
ggplot(
    follow_up_messages_count %>%
      select(val, stat) %>%
      group_by(stat) %>%
      summarize(val = sum(val)),
    aes(x = stat, y = val)
  ) + geom_bar(stat = 'identity') + ggtitle("# of Follow-Up Messages, Across Utterance Type") +
    xlab('Message Type') + ylab('# of Messages Produced')
```

```{r fig.width=7, fig.height=4, echo=FALSE}
ggplot(
    follow_up_messages_count,
    aes(x = reorder(role, desc(role)), y = val, fill=stat)
  ) + geom_bar(stat = 'identity', position = 'stack') + ggtitle("# of Follow-Up Messages, Across Utterance Type") +
    xlab('Utterance Type') + ylab('# of Messages Produced')
```

## Quantifier Specific Language
```{r}
quantifier_messages_count = chat_messages %>%
  filter(
    utteranceType == 'Quantifier' | 
      (utteranceType == 'Question' & 
         (grepl('any', text)  | 
            grepl('all', text) |
            grepl('mostly', text) | 
            grepl('some', text)
          )
       )
  ) %>%
  mutate(utteranceTypeSignifiers=replace(utteranceTypeSignifiers, utteranceType=='Question', 'Question')) %>%
  select(role, utteranceTypeSignifiers) %>%
  group_by(role, utteranceTypeSignifiers) %>%
  summarize(
    Count = length(utteranceTypeSignifiers)
  )
```

```{r fig.width=7, fig.height=4, echo=FALSE}
ggplot(
    quantifier_messages_count,
    aes(x = utteranceTypeSignifiers, y = Count)
  ) + geom_bar(stat = 'identity') + ggtitle("# of Messages with Quantifier Signifier (Teacher & Student Combined)") +
    xlab('Signifier') + ylab('# of Times Utilized')
```


```{r fig.width=7, fig.height=4, echo=FALSE}
ggplot(
    quantifier_messages_count,
    aes(x = role, y = Count, fill=utteranceTypeSignifiers)
  ) + geom_bar(stat = 'identity', position ='stack') + ggtitle("# of Messages with Quantifier Signifier (Teacher & Student Split)") +
    xlab('Role') + ylab('# of Times Utilized')
```

## Analysis of Langauge Across Pairs
```{r}
game_informative_msgs = chat_messages %>%
  filter(messageType == 'Informative') %>%
  group_by(gameid) %>%
  select(gameid, utteranceType) %>%
  summarize(
    None = sum(utteranceType == 'None'),
    Quantifier = sum(utteranceType == 'Quantifier'),
    Conditional = sum(utteranceType == 'Conditional'),
    Exemplar = sum(utteranceType == 'Exemplar'),
    Generic = sum(utteranceType == 'Generic'),
    Imperative = sum(utteranceType == 'Imperative'),
    Affirmative = sum(utteranceType == 'Affirmative'),
    Knowledge = sum(utteranceType == 'Knowledge'),
    Logic = sum(utteranceType == 'Logic'),
    Question = sum(utteranceType == 'Question'),
    Possibility_Conditional = sum(utteranceType == 'Possibility,Conditional'),
  ) %>%
  select(gameid, None, Quantifier, Conditional, Exemplar, Generic, Imperative, Affirmative, Knowledge, Logic, Question, Possibility_Conditional) %>%
  gather(key = "stat", value = "val", -gameid)
```

```{r fig.width=7, fig.height=4, echo=FALSE}
ggplot(
    game_informative_msgs,
    aes(x='', y=val, fill=stat)
  ) + geom_bar(stat = 'identity', position ='stack') + ggtitle("Distribution of Utterance Types (Informative) Across Teachers") +
    xlab('Teacher') + ylab('# of Messages of Utterance Type') + facet_grid(~gameid)
```

```{r}
pair_messages = chat_messages %>%
  filter(messageType == 'Follow-Up' | messageType == 'Informative') %>%
  mutate(role = as.character(role)) %>%
  mutate(role = replace(role, role == "explorer", "T")) %>%
  mutate(role = replace(role, role == "student", "S")) %>%
  group_by(gameid, role) %>%
  select(role, gameid, utteranceType) %>%
  summarize(
    None = sum(utteranceType == 'None'),
    Quantifier = sum(utteranceType == 'Quantifier'),
    Conditional = sum(utteranceType == 'Conditional'),
    Exemplar = sum(utteranceType == 'Exemplar'),
    Generic = sum(utteranceType == 'Generic'),
    Imperative = sum(utteranceType == 'Imperative'),
    Affirmative = sum(utteranceType == 'Affirmative'),
    Knowledge = sum(utteranceType == 'Knowledge'),
    Logic = sum(utteranceType == 'Logic'),
    Question = sum(utteranceType == 'Question'),
    Possibility_Conditional = sum(utteranceType == 'Possibility,Conditional'),
  ) %>%
  select(role, gameid, None, Quantifier, Conditional, Exemplar, Generic, Imperative, Affirmative, Knowledge, Logic, Question, Possibility_Conditional) %>%
  gather(key = "stat", value = "val", 3:13)
```

```{r fig.width=7, fig.height=4, echo=FALSE}
ggplot(
    pair_messages,
    aes(x=role, y=val, fill=stat)
  ) + geom_bar(stat = 'identity', position ='stack') + ggtitle("Distribution of Utterance Types (Informative + Follow-Up) Across Player Pairs ") +
    xlab('Role') + ylab('# of Messages of Utterance Type') + facet_grid(~gameid)
```

```{r}
ggplot(y,
    aes(x = rule_idx, y = n)
) + geom_bar(stat = 'identity') + ggtitle("# of Games By Rule Index") +
    xlab('Rule Index') + ylab('# Games Played')
```

## Analysis of Language Across Concepts

## Analysis of Language Across Concept Types 


